\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{verbatim}

\title{A Public End-to-End Full-Key Recovery Attack on the \texttt{main.c} Implementation of SEPAR}
\author{}
\date{2026-04-14}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{proposition}{Proposition}

\newcommand{\Z}[1]{\mathbb{Z}_{#1}}
\newcommand{\hi}{\operatorname{hi}}
\newcommand{\lo}{\operatorname{lo}}
\newcommand{\rot}{\operatorname{rotl}_{16}}
\newcommand{\can}{\operatorname{can}}

\begin{document}
\maketitle

\begin{abstract}
We study the \emph{implementation} of SEPAR in \texttt{SEPAR/main.c}, not the abstract cipher description. In the chosen-IV, chosen-plaintext, chosen-ciphertext, reset model, we give a public end-to-end full-key recovery attack against the 256-bit implementation key. The attack is exact once a suitable IV is found. Its only stochastic component is the chosen-IV search. The deterministic recovery pipeline has three parts: an exact stage-8 public bootstrap; an exact same-stage inward recursion that recovers $(K_7,s_8),\dots,(K_2,s_3)$; and an exact stage-1 closure from the recovered stage-2 reduced table. The final acceptance condition is exact: a candidate 256-bit key is accepted if and only if it reproduces the full one-word oracle table on the successful IV. We give exact mathematical definitions for every primitive used by the implementation, state the full attack as a sequence of explicit public transformations on $2^{16}$-entry tables, prove conditional correctness, report a complete random-key recovery, and give a Monte Carlo estimate for the success probability of the IV-search phase.
\end{abstract}

\section{Introduction}
The target of this paper is the concrete word-oriented implementation in \texttt{SEPAR/main.c}. The relevant attack surface is not the abstract design claim; it is the resettable 16-bit one-word interface induced by the code. Once an IV is fixed, the implementation exposes an exact public permutation of $\Z{2^{16}}$ on one-word inputs. Because the block size is only 16 bits, one can recover and manipulate the full codebook exactly.

The attack model is:
\begin{itemize}[leftmargin=1.5em]
\item chosen IV;
\item chosen plaintext and chosen ciphertext;
\item exact state reset between oracle calls.
\end{itemize}

Under that model, each IV induces a public permutation
\[
F_{\mathrm{IV}}:\Z{2^{16}}\to\Z{2^{16}}
\]
that can be recovered in full with $2^{16}$ chosen encryptions. The inverse table can be recovered either by chosen-ciphertext queries or, as in the implementation used here, by inverting the public encryption table locally.

\paragraph{Main claim.}
For the tested random key
\[
K=
\texttt{2DE13EF0284A5C9BEB9FA9664ED8258BBDE5A7EA23C03CE6F3DD038E0E7570B5},
\]
the public attack implemented in
\[
\texttt{C:/Users/noaho/Desktop/upload/separ\_public\_full\_hybrid\_attack.c}
\]
recovers the full 256-bit key by:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item a chosen-IV beam search and exact hybrid public bootstrap for $K_8$;
\item on a successful inward IV, the same exact stage-local recursion for $(K_7,s_8),\dots,(K_2,s_3)$;
\item an exact stage-1 closure from the public stage-2 reduced table;
\item exact full-codebook verification of the recovered candidate key.
\end{enumerate}

What is proved here is therefore conditional exactness: once the chosen-IV search hits a suitable outer bootstrap IV and a suitable inward recursive IV, the rest of the attack is deterministic and exact.

\section{Exact implementation model}
\subsection{Stage-local block permutation}
Let the eight 32-bit stage keys be
\[
K_i=(k_{i,0},k_{i,1})\in \Z{2^{16}}\times \Z{2^{16}},
\qquad 1\le i\le 8.
\]

Let $\Sigma:\Z{2^{16}}\to\Z{2^{16}}$ be the nibblewise S-box layer induced by the fixed tables $S_1,S_2,S_3,S_4$ in the implementation, and let $\Sigma^{-1}$ be the inverse layer induced by $\texttt{IS1},\dots,\texttt{IS4}$.

Let $L:\Z{2^{16}}\to\Z{2^{16}}$ be the exact mixing layer implemented by \texttt{sep\_rotl16}. If
\[
x=(a,b,c,d)\in (\Z{16})^4
\]
in nibble notation, define
\[
M(a,b,c,d)=(a\oplus c,\, b\oplus d,\, c\oplus b,\, d\oplus a),
\]
identify nibble tuples with 16-bit words, and set
\[
L(x)=M(x)\oplus \rot(M(x),12)\oplus \rot(M(x),8).
\]
Its inverse is the exact \texttt{sep\_inrotl16} routine.

For each stage $i$, the code derives two additional 16-bit subkeys $\kappa_{i,2},\kappa_{i,3}$ from $(k_{i,0},k_{i,1})$:
\begin{align*}
u_i &= \rot(k_{i,0},6),&
v_i &= \rot(k_{i,1},10),\\
\kappa_{i,2} &= \bigl(u_i \vee (S_1((u_i \gg 6)\bmod 16)\ll 6)\bigr)\oplus (i+2),\\
\kappa_{i,3} &= \bigl(v_i \vee (S_1((v_i \gg 6)\bmod 16)\ll 6)\bigr)\oplus (i+3),
\end{align*}
where $\vee$ is bitwise OR and all arithmetic is in the exact 16-bit sense used by the implementation.

Define the XOR translation $A_t(x)=x\oplus t$. Then the stage-local encryption permutation implemented by \texttt{enc\_block} is exactly
\begin{equation}
\label{eq:stage-enc}
E_i
=
A_{\kappa_{i,2}\oplus \kappa_{i,3}}
\circ \Sigma
\circ A_{k_{i,0}\oplus k_{i,1}}
\circ L
\circ \Sigma
\circ A_{\kappa_{i,3}}
\circ L
\circ \Sigma
\circ A_{\kappa_{i,2}}
\circ L
\circ \Sigma
\circ A_{k_{i,1}}
\circ L
\circ \Sigma
\circ A_{k_{i,0}},
\end{equation}
and $D_i=E_i^{-1}$ is the exact \texttt{dec\_block} inverse.

\subsection{Translated stage action and full one-word encryption}
For a state word $s_i\in\Z{2^{16}}$, define the translated stage permutation
\[
B_i^{(s_i)}(x)=E_i(x+s_i),
\qquad
\bigl(B_i^{(s_i)}\bigr)^{-1}(y)=D_i(y)-s_i,
\]
where $+$ and $-$ are addition and subtraction in $\Z{2^{16}}$.

For a fixed reset IV with state $(s_1,\dots,s_8)$, the public one-word encryption permutation is
\[
F_{\mathrm{IV}}
=
B_8^{(s_8)}
\circ B_7^{(s_7)}
\circ \cdots \circ
B_1^{(s_1)}.
\]

\subsection{State update law}
Write the eight state words as $(s_1,\dots,s_8)$ and the stage outputs as
\[
v_{12},v_{23},v_{34},v_{45},v_{56},v_{67},v_{78},c.
\]
From \texttt{Separ\_Encryption} in \texttt{main.c}, the exact state update after one word is
\begin{align}
s_2' &= s_2 + v_{12} + v_{56} + s_6, \label{eq:s2}\\
s_3' &= s_3 + v_{23} + v_{34} + s_4 + s_1, \label{eq:s3}\\
s_4' &= s_4 + v_{12} + v_{45} + s_8, \label{eq:s4}\\
s_5' &= s_5 + v_{23} + L', \label{eq:s5}\\
s_6' &= s_6 + v_{12} + v_{45} + s_7, \label{eq:s6}\\
s_7' &= s_7 + v_{23} + v_{67}, \label{eq:s7}\\
s_8' &= s_8 + v_{45}, \label{eq:s8}\\
s_1' &= s_1 + s_5 + v_{34} + 2v_{23} + v_{78}. \label{eq:s1}
\end{align}
The attack below uses only reset one-word tables, so these equations matter only because they determine which state word appears as the unknown output translation at each recursive step.

\subsection{Oracle model}
Under chosen IV and reset, each IV defines exact public codebooks
\[
F_{\mathrm{IV}}(x)=\operatorname{Enc}_{K,\mathrm{IV}}(x),
\qquad
F_{\mathrm{IV}}^{-1}(y)=\operatorname{Dec}_{K,\mathrm{IV}}(y),
\]
and the implementation used here obtains the inverse table locally by inverting the public encryption table.

\section{Exact public tables used by the attack}
\subsection{Recursive current tables}
For $m\in\{1,\dots,7\}$ define the inner cascade
\[
R_m
=
B_m^{(s_m)}
\circ B_{m-1}^{(s_{m-1})}
\circ \cdots \circ
B_1^{(s_1)}.
\]

\begin{definition}[Current table at level $m$]
Assume $K_{m+1},\dots,K_8$ and $s_{m+2},\dots,s_8$ are already known for the fixed IV. The \emph{current table} at level $m$ is the public table
\[
C_m:\Z{2^{16}}\to\Z{2^{16}}
\]
computed recursively by
\[
C_7=D_8\circ F_{\mathrm{IV}},
\qquad
C_m=D_{m+1}\bigl(C_{m+1}-s_{m+2}\bigr)
\quad (1\le m\le 6).
\]
\end{definition}

\begin{lemma}[Exact translation form of the current table]
\label{lem:current-table}
For every $m\in\{1,\dots,7\}$,
\[
C_m(x)=R_m(x)+s_{m+1}
\qquad\text{for all }x\in\Z{2^{16}}.
\]
\end{lemma}

\begin{proof}
For $m=7$,
\[
F_{\mathrm{IV}}(x)=E_8(R_7(x)+s_8),
\]
hence
\[
C_7(x)=D_8(F_{\mathrm{IV}}(x))=R_7(x)+s_8.
\]
Assume the statement holds at level $m+1$. Then
\[
C_{m+1}(x)=R_{m+1}(x)+s_{m+2}
=
E_{m+1}(R_m(x)+s_{m+1})+s_{m+2},
\]
so
\[
C_m(x)
=
D_{m+1}(C_{m+1}(x)-s_{m+2})
=
D_{m+1}(E_{m+1}(R_m(x)+s_{m+1}))
=
R_m(x)+s_{m+1}.
\]
\end{proof}

Thus the exact public object attacked at level $m$ is not an arbitrary hidden function. It is the inner cascade $R_m$ observed through a single output translation by $s_{m+1}$.

\subsection{Exact low-byte score}
Let
\[
H(y)=\left\lfloor \frac{y}{2^8}\right\rfloor \in \Z{256}
\]
be the output high byte. For a candidate low byte $b\in\Z{256}$ and a row index $r\in\Z{256}$ define
\[
\sigma_{m,b}(r)
=
\left|
\left\{
H\bigl(C_m(256r+\lambda)-b\bigr)
:
\lambda\in\Z{256}
\right\}
\right|,
\]
and the exact total score
\[
\Sigma_m(b)=\sum_{r=0}^{255}\sigma_{m,b}(r).
\]
This is exactly the quantity computed by \texttt{exact\_low\_byte\_scan\_u16()}. The implementation evaluates all $256$ low-byte hypotheses against all $256$ rows and all $256$ row entries.

\subsection{Projected quotient cycles}
Define the 4-bit projection
\[
\pi(y)=H(y)\bmod 16 \in \Z{16}.
\]
For a stage key pair $K_i$ define the stage-$i$ quotient map
\[
q_{i,K_i}(h)=\pi(E_i(256h)),
\qquad h\in\Z{16},
\]
and the full high-byte quotient map
\[
Q_{i,K_i}(h)=H(E_i(256h)),
\qquad h\in\Z{256}.
\]

The implementation uses the fixed exact auxiliary sets
\[
\Delta_0=\{1,2,4,8,15,16\},
\]
\[
\Omega=\{0,1,2,3,5,8,13,21,34,55,89,144,200,233,255\},
\]
and
\[
\mathcal R=\{0x00,0x55,0xAA,0xFF\}.
\]

\subsection{Representative tables and invisible bits}
The code does not enumerate all $2^{32}$ stage pairs. It uses exact stage-specific invisible-bit sets. For stages $2,\dots,8$, the hard-coded sets are
\begin{align*}
P_2=P_3
&=\{0,1,6,7,12,13,14,15,18,19,20,21,22,23\},\\
P_4
&=\{6,7,12,13,14,15,18,19,20,21,22,23\},\\
P_5
&=\{0,6,7,12,13,14,15,18,19,20,21,22,23\},\\
P_6=P_7
&=\{0,1,2,6,7,12,13,14,15,18,19,20,21,22,23\},\\
P_8
&=\{0,1,6,7,12,13,14,15,18,19,20,21,22,23,29\}.
\end{align*}
Let
\[
V_i=\{0,\dots,31\}\setminus P_i.
\]
Then
\[
|V_2|=|V_3|=18,\quad
|V_4|=20,\quad
|V_5|=19,\quad
|V_6|=|V_7|=|V_8|=17.
\]

For each stage $i$, the representative table enumerates all visible masks in $\{0,1\}^{V_i}$ with all invisible bits zero, computes the corresponding quotient cycle $q_{i,K_i}$, and groups visible assignments by that cycle. For stages $2,\dots,7$, the cycle is canonicalized by rotation so that the first entry is $0$; stage~8 uses raw cycles.

\subsection{Cycle edge score and exact order reconstruction}
For any public table $T:\Z{2^{16}}\to\Z{2^{16}}$, define the exact projected edge-count matrix
\[
M_T(a,b)
=
\left|
\left\{
(x,\delta)\in \Z{2^{16}}\times \Delta_0:
\pi(T(x))=a,\;
\pi(T(x+\delta))=b
\right\}
\right|,
\qquad a,b\in\Z{16}.
\]
If $c=(c_0,\dots,c_{15})$ is a 16-cycle, define its exact edge score
\[
S_T(c)=\sum_{j=0}^{15} M_T(c_j,c_{j+1\bmod 16}).
\]

The cycle score recovers only the 16-point projected cycle. To lift it to a full 256-point high-byte ordering, the implementation builds the exact 8-bit transition matrix
\[
N_T(u,v)
=
\left|
\left\{
(x,\delta)\in \Z{2^{16}}\times \Delta_0:
H(T(x))=u,\;
H(T(x+\delta))=v
\right\}
\right|,
\qquad u,v\in\Z{256},
\]
and then, for each adjacent pair of cycle layers, solves an exact $16\times 16$ maximum-weight assignment by the Hungarian algorithm. The resulting wrap permutation is accepted only if it forms one 16-cycle on the path labels. This yields the exact order array used by the candidate lift.

\subsection{Exact candidate lift and verifier}
For any recovered cycle/order pair, the implementation:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item enumerates every representative-group base key compatible with the cycle;
\item expands all invisible masks in $P_i$;
\item computes the full quotient bytes $Q_{i,K_i}$;
\item tests exact equality against the reconstructed order targets first on the necessary sample set $\Omega$, then on all $256$ values;
\item for stages $2,\dots,7$, sets
\[
s_{i+1}=\lo(s_{i+1}) + 256\tau
\]
from the recovered byte translation $\tau$;
\item exact-scores the surviving candidates by
\[
V_i(K_i;T)=\sum_{r\in\mathcal R}
\min_{b\in\Z{256}}
\left|
\left\{
H\bigl(D_i(T(256r+\lambda))-b\bigr):\lambda\in\Z{256}
\right\}
\right|.
\]
\end{enumerate}
The sample test is not heuristic: full equality on $\Z{256}$ implies equality on $\Omega$, so the sample check discards only impossible candidates.

\section{The outer public bootstrap for \texorpdfstring{$K_8$}{K8}}
\subsection{Weak-IV beam search}
For an IV, define the exact row score
\[
\rho(\mathrm{IV})
=
\left|
\left\{
H(F_{\mathrm{IV}}(\lambda)) : \lambda\in\Z{256}
\right\}
\right|.
\]
The implementation keeps the $B$ IVs with smallest $\rho(\mathrm{IV})$. This is a search-ordering device only. It is not part of the acceptance criterion.

\subsection{Hybrid exact bootstrap}
For each IV in the beam, the implementation:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item computes the exact stage-8 quotient/bootstrap candidate set by projected-cycle maximization and exact verifier scoring;
\item closes that set under the independent MSB toggles
\[
(a,b)\mapsto (a\oplus \epsilon_0 2^{15},\, b\oplus \epsilon_1 2^{15});
\]
\item exact-scores the resulting small pool by the additive-differential score
\[
A_{\mathrm{IV}}(K_8)
=
\sum_{\delta\in\Delta_0}
\max_{\Delta\in\Z{2^{16}}}
\left|
\left\{
x : D_8(F_{\mathrm{IV}}(x+\delta);K_8)-D_8(F_{\mathrm{IV}}(x);K_8)=\Delta
\right\}
\right|;
\]
\item aggregates those exact scores across the beam and outputs the best $K_8$.
\end{enumerate}

For the tested random key, the 512-trial beam with beam size $16$ recovered the correct
\[
K_8=(\texttt{0E75},\texttt{70B5})
\]
on search seeds $1,2,3$. This is empirical evidence, not a proof of key-generic stage-8 success.

\section{Exact same-stage recursion from \texorpdfstring{$K_7$}{K7} to \texorpdfstring{$K_2$}{K2}}
\subsection{Deterministic recursion once \texorpdfstring{$K_8$}{K8} is known}
Fix an IV and assume the outer bootstrap has already recovered the correct $K_8$. The inward recursion then repeatedly applies the exact stage-local scan to the current tables of Lemma~\ref{lem:current-table}.

For $m=7,6,5,4,3,2$:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item compute the current table $C_m$;
\item compute the exact low-byte score $\Sigma_m$ and let $\ell_m$ be its minimizing low byte;
\item form the corrected table
\[
\widetilde{C}_m(x)=C_m(x)-\ell_m;
\]
\item compute all maximizing projected cycles for $\widetilde{C}_m$;
\item reconstruct all exact compatible orders;
\item lift to the full candidate family $(K_m,s_{m+1})$;
\item exact-score all candidates by the verifier $V_m$ and output the lexicographic winner;
\item peel the recovered stage:
\[
C_{m-1}=D_m(C_m-s_{m+1};K_m).
\]
\end{enumerate}

\begin{definition}[Recursive IV]
An IV is \emph{recursive to depth $d$} if, when the above deterministic procedure is run with the correct $K_8$, it returns the true pairs/states for stages
\[
7,6,\dots,d.
\]
An IV is \emph{fully inward-recursive} if it is recursive to depth $2$.
\end{definition}

\paragraph{Why the recursion stops at stage~2.}
The projected-family scan exists only for stages $2,\dots,8$. Stage~1 has no next visible stage and therefore no stage-1 analogue of the quotient cycle. Stage~1 must be closed separately from the stage-2 reduced table.

\section{Exact stage-1 closure from the public stage-2 table}
\subsection{Reduction to a single translated stage}
Assume the inward recursion has recovered the true $(K_2,s_3)$ on a fully inward-recursive IV. Then
\[
C_2(x)=E_2(E_1(x+s_1)+s_2)+s_3.
\]
Define
\[
W_2(x)=D_2(C_2(x)-s_3;K_2).
\]
Then exactly
\begin{equation}
\label{eq:stage1-reduction}
W_2(x)=E_1(x+s_1)+s_2.
\end{equation}

\subsection{Low-byte recovery for \texorpdfstring{$s_2$}{s2}}
The implementation first applies the same exact low-byte scan to $W_2$. For each candidate
\[
s_2^\star\in \Z{2^{16}},
\]
it forms
\[
C^\star(x)=W_2(x)-s_2^\star.
\]
Any true stage-1 solution must satisfy
\[
C^\star(x)=E_1(x+s_1;K_1).
\]

\subsection{Step 1: exact row-signature inversion}
For $u\in\Z{256}$ define the 16-point row signature
\[
R_u^{C^\star}(h)=\pi(C^\star(u+256h)),
\qquad h\in\Z{16}.
\]
Write
\[
s_1 = s_{1,\mathrm{lo}} + 256\,s_{1,\mathrm{mid}} + 4096\,s_{1,\mathrm{hi}},
\]
with $s_{1,\mathrm{lo}}\in\Z{256}$ and $s_{1,\mathrm{mid}},s_{1,\mathrm{hi}}\in\Z{16}$.

For a candidate $s_{1,\mathrm{lo}}$, the routine \texttt{stage1\_step1\_candidates\_u16()} checks the exact carry partition:
\begin{quote}
all rows with $u+s_{1,\mathrm{lo}}<256$ must share one common signature $\sigma$, and all rows with $u+s_{1,\mathrm{lo}}\ge 256$ must share the one-step rotation $\operatorname{rot}_1(\sigma)$.
\end{quote}
For each surviving $s_{1,\mathrm{lo}}$ and each $s_{1,\mathrm{mid}}\in\Z{16}$, the implementation rotates $\sigma$ by $16-s_{1,\mathrm{mid}}$ and inverts it through the exact precomputed family
\[
\nu_{a,b,c,d}:\Z{16}\to\Z{16},
\]
where
\[
\nu_{a,b,c,d}(x)=S_2\bigl(t_4(x)\oplus a\oplus b\bigr)\oplus c\oplus d
\]
with
\begin{align*}
t_1(x)&=S_2(x\oplus a),\\
t_2(x)&=S_2(t_1(x)\oplus b),\\
t_3(x)&=S_2(t_2(x)\oplus c),\\
t_4(x)&=S_2(t_3(x)\oplus d).
\end{align*}
This yields exact candidates for
\[
s_{1,\mathrm{lo}},\quad s_{1,\mathrm{mid}},\quad a,\quad b,\quad c,\quad d.
\]

\subsection{Step 2: exact lift to full \texorpdfstring{$K_1$}{K1} candidates}
The routine next uses exact inverse-image tables for the derived-key high nibbles of $\kappa_{1,2}$ and $\kappa_{1,3}$. For each surviving $(s_{1,\mathrm{lo}},s_{1,\mathrm{mid}},a,b,c,d)$, each
\[
\ell_6\in I_2(c),\qquad \beta\in I_3(d),
\]
and each 12-bit free mask $f\in\Z{2^{12}}$, it constructs
\begin{align}
k_{1,0} &=
\ell_6
\;\vee\;
\bigl((f\bmod 4)\ll 6\bigr)
\;\vee\;
(a\ll 8)
\;\vee\;
\left(\left(\left\lfloor\frac{f}{4}\right\rfloor \bmod 16\right)\ll 12\right), \label{eq:k10-lift}\\
k_{1,1} &=
(\beta\bmod 4)
\;\vee\;
\left(\left(\left\lfloor\frac{f}{64}\right\rfloor \bmod 64\right)\ll 2\right)
\;\vee\;
(b\ll 8)
\;\vee\;
\left(\left(\left\lfloor\frac{\beta}{4}\right\rfloor \bmod 16\right)\ll 12\right). \label{eq:k11-lift}
\end{align}
This is the exact bit lift implemented by \texttt{stage1\_step2\_candidates\_u16()}.

The resulting pair is exact-tested on the 12-bit slice
\[
X_{12}=\{0,\dots,4095\}
\]
by comparing nibble $(7..4)$ values:
\[
\left\lfloor \frac{E_1((x+s_{1,\mathrm{lo}}+256s_{1,\mathrm{mid}})\bmod 2^{12};K_1)}{2^4}\right\rfloor \bmod 16
=
\left\lfloor \frac{C^\star(x)}{2^4}\right\rfloor \bmod 16
\quad\text{for all }x\in X_{12}.
\]

\subsection{Step 3: exact recovery of the top nibble of \texorpdfstring{$s_1$}{s1}}
For each surviving pair $K_1$, the implementation scans all 16 possibilities for $s_{1,\mathrm{hi}}$ and performs full exact verification:
\[
E_1(x+s_1;K_1)=C^\star(x)
\qquad\text{for all }x\in\Z{2^{16}}.
\]
If exactly one triple $(K_1,s_1,s_2)$ survives, the stage-1 closure succeeds.

\begin{theorem}[Exactness of the stage-1 closure]
\label{thm:stage1}
Let $W_2$ be the exact public table of~\eqref{eq:stage1-reduction}. If there is a unique triple $(K_1,s_1,s_2)$ satisfying
\[
W_2(x)=E_1(x+s_1;K_1)+s_2
\qquad\text{for all }x\in\Z{2^{16}},
\]
then \texttt{stage1\_finish\_from\_stage2\_current()} returns exactly that triple.
\end{theorem}

\begin{proof}
The routine is exhaustive over all candidates for the low and high bytes of $s_2$, all exact row-signature solutions for $(s_{1,\mathrm{lo}},s_{1,\mathrm{mid}},a,b,c,d)$, all inverse-image candidates from $I_2(c)$ and $I_3(d)$, all $2^{12}$ free masks in~\eqref{eq:k10-lift}--\eqref{eq:k11-lift}, and all 16 possibilities for $s_{1,\mathrm{hi}}$. Every surviving candidate is tested by exact equality either on the 12-bit slice or on the full $2^{16}$-entry table, and the routine accepts only if the number of full solutions is exactly one. Hence a unique solution is returned exactly.
\end{proof}

\section{Public full-key acceptance and conditional correctness}
\begin{definition}[Exact public acceptance test]
Given recovered pairs
\[
\widehat{K}=(\widehat{K}_1,\dots,\widehat{K}_8),
\]
rebuild the candidate one-word table
\[
\widehat{F}_{\mathrm{IV}}
\]
under the same IV and accept if and only if
\[
\widehat{F}_{\mathrm{IV}}=F_{\mathrm{IV}}
\]
as functions on $\Z{2^{16}}$.
\end{definition}

Because the block size is only 16 bits, this is an exact comparison of two $65536$-entry tables.

\begin{definition}[Outer-success and inward-success events]
Fix the exact IV sequence tested by the implementation.
\begin{enumerate}[label=(\alph*),leftmargin=1.5em]
\item The \emph{outer-success event} is the event that the beam search and exact hybrid aggregate rank the true $K_8$ first.
\item The \emph{inward-success event} for an IV is the event that the same-stage recursion returns the true $(K_7,s_8),\dots,(K_2,s_3)$ and the stage-1 solver returns a unique solution.
\end{enumerate}
\end{definition}

\begin{theorem}[Conditional exactness of the full public attack]
\label{thm:main}
Assume the outer-success event occurs on one of the tested beam IVs, and assume one of the tested inward IVs satisfies the inward-success event. Then the implementation outputs the true 256-bit key and accepts it by the exact codebook test. No false full key can be accepted on that IV.
\end{theorem}

\begin{proof}
Under outer-success, the recovered $K_8$ is the true $K_8$. By Lemma~\ref{lem:current-table}, each inward recursion level is then applied to the exact public current table $C_m=R_m+s_{m+1}$. Under inward-success, the recursion returns the true pairs/states down to $(K_2,s_3)$. Equation~\eqref{eq:stage1-reduction} therefore holds for the true $W_2$, and Theorem~\ref{thm:stage1} returns the unique true stage-1 triple $(K_1,s_1,s_2)$. The reconstructed 256-bit key is thus the true implementation key. The final acceptance test compares full tables, so no false key can satisfy it on the successful IV.
\end{proof}

\section{Worked random-key recovery}
\subsection{Target key and successful IVs}
The tested random key is
\[
K=
\texttt{2DE13EF0284A5C9BEB9FA9664ED8258BBDE5A7EA23C03CE6F3DD038E0E7570B5}.
\]

The successful public run used:
\begin{itemize}[leftmargin=1.5em]
\item outer beam budget $N_{\mathrm{out}}=512$;
\item beam size $B=16$;
\item search seed $1$;
\item successful inward IV
\[
\texttt{D17A5783A9F94E08F948B6EB536519A4}.
\]
\end{itemize}

\subsection{Recovered values}
The public recovery returned
\begin{align*}
K_8 &= (\texttt{0E75},\texttt{70B5}),\\
K_7 &= (\texttt{F3DD},\texttt{038E}),\qquad s_8=\texttt{BE8C},\\
K_6 &= (\texttt{23C0},\texttt{3CE6}),\qquad s_7=\texttt{E981},\\
K_5 &= (\texttt{BDE5},\texttt{A7EA}),\qquad s_6=\texttt{EA65},\\
K_4 &= (\texttt{4ED8},\texttt{258B}),\qquad s_5=\texttt{44A8},\\
K_3 &= (\texttt{EB9F},\texttt{A966}),\qquad s_4=\texttt{E26E},\\
K_2 &= (\texttt{284A},\texttt{5C9B}),\qquad s_3=\texttt{BA5B},\\
K_1 &= (\texttt{2DE1},\texttt{3EF0}),\qquad s_2=\texttt{461F},\qquad s_1=\texttt{DF21}.
\end{align*}

The decisive public log line is
\begin{verbatim}
full-key candidate verified against exact one-word oracle table
\end{verbatim}
from \texttt{separ\_public\_full\_hybrid\_attack\_clean\_verbose.log}.

\section{Monte Carlo estimate for IV-search success}
\subsection{Deterministic core versus stochastic front-end}
The attack splits into:
\begin{itemize}[leftmargin=1.5em]
\item a deterministic algebraic core: once a suitable IV is fixed, all recovery and verification steps are exact;
\item a stochastic IV-search front-end: the attack samples IVs until outer-success and inward-success occur.
\end{itemize}

For a fixed recovered $K_8$, model the inward search as a Bernoulli process with success probability
\[
p_{\mathrm{in}}
=
\Pr[\text{a random IV is fully inward-recursive}\mid K_8 \text{ already recovered} ].
\]

\subsection{Pilot estimate}
Pooling the earlier inward-recursion pilot scans with the first observed full success gives
\[
x=1 \text{ success in } n=1727 \text{ random inward IV trials},
\]
hence
\[
\widehat{p}_{\mathrm{in}}=\frac{1}{1727}\approx 5.79\times 10^{-4}.
\]
The Wilson 95\% interval is approximately
\[
0.000102 \lesssim p_{\mathrm{in}} \lesssim 0.003273.
\]
These values are empirical measurements for the tested random-key instance, not a theorem about arbitrary keys.

\subsection{Budget-to-success conversion}
For an inward IV budget of $N$ independent random trials,
\[
P_{\mathrm{succ}}(N)=1-(1-\widehat{p}_{\mathrm{in}})^N.
\]
The corresponding point estimates are:
\begin{center}
\begin{tabular}{rrrr}
\toprule
Inward IV budget $N$ & Point estimate & Wilson low & Wilson high \\
\midrule
256   & 0.138 & 0.026 & 0.568 \\
512   & 0.257 & 0.051 & 0.813 \\
1024  & 0.447 & 0.099 & 0.965 \\
2048  & 0.695 & 0.189 & 0.999 \\
4096  & 0.907 & 0.342 & 1.000 \\
8192  & 0.991 & 0.567 & 1.000 \\
10000 & 0.997 & 0.640 & 1.000 \\
\bottomrule
\end{tabular}
\end{center}

\section{Complexity and bound}
\subsection{Oracle-query complexity}
One full one-word codebook requires exactly
\[
2^{16}
\]
chosen encryptions. The implementation does not require chosen-ciphertext queries, so
\[
Q_{\mathrm{dec}}=0.
\]

Suppose the attack is run with:
\begin{itemize}[leftmargin=1.5em]
\item outer IV-search budget $N_{\mathrm{out}}$;
\item beam size $B$;
\item one distinguished initial inward IV;
\item inward random-IV budget $N_{\mathrm{in}}$.
\end{itemize}
Then the worst-case number of codebooks built by the current implementation is
\[
N_{\mathrm{out}} + B + 1 + B + N_{\mathrm{in}}
=
N_{\mathrm{out}} + 2B + N_{\mathrm{in}} + 1.
\]
Hence the exact chosen-encryption query bound is
\[
Q_{\mathrm{enc}}
=
\bigl(N_{\mathrm{out}} + 2B + N_{\mathrm{in}} + 1\bigr)2^{16}.
\]

For the validated successful run reported above,
\[
N_{\mathrm{out}}=512,\qquad B=16,\qquad N_{\mathrm{in}}=1,
\]
and success occurred on the distinguished initial inward IV. The realized query count was therefore
\[
(512+16+1)\cdot 2^{16}
=
529\cdot 65536
=
34{,}668{,}544.
\]

\subsection{Exact local work}
Once a codebook is built, all remaining work is local and exact. The fixed representative-table sizes used by the implementation are:
\begin{center}
\begin{tabular}{rcc}
\toprule
Stage & $|V_i|$ & Representative masks \\
\midrule
8 & 17 & $2^{17}=131072$ \\
7 & 17 & $2^{17}=131072$ \\
6 & 17 & $2^{17}=131072$ \\
5 & 19 & $2^{19}=524288$ \\
4 & 20 & $2^{20}=1048576$ \\
3 & 18 & $2^{18}=262144$ \\
2 & 18 & $2^{18}=262144$ \\
\bottomrule
\end{tabular}
\end{center}

On the successful inward IV, the cycle-search logs recorded:
\begin{center}
\begin{tabular}{rcc}
\toprule
Stage & Representative/cycle space & Logged cycle work \\
\midrule
7 & $2^{17}$ reps, shifted cycle search & $2^{19}=524288$ \\
6 & $2^{17}$ reps, shifted cycle search & $2^{19}=524288$ \\
5 & $2^{19}$ reps, shifted cycle search & $2^{20}=1048576$ \\
4 & $2^{20}$ reps, shifted cycle search & $2^{20}=1048576$ \\
3 & $2^{18}$ reps, shifted cycle search & $3\cdot 2^{18}=786432$ \\
2 & $2^{18}$ reps, shifted cycle search & $3\cdot 2^{18}=786432$ \\
\bottomrule
\end{tabular}
\end{center}

The stage-1 closure then adds:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item an exact low-byte scan for $s_2$;
\item all $256$ high-byte hypotheses for $s_2$;
\item exact inversion of the row-signature predicate;
\item exact lift over the remaining $2^{12}$ free key bits;
\item full table verification of the surviving stage-1 candidates.
\end{enumerate}

\subsection{Overall bound}
For fixed budgets $(N_{\mathrm{out}},B,N_{\mathrm{in}})$, the implemented attack has exact query complexity
\[
Q_{\mathrm{enc}}
=
\bigl(N_{\mathrm{out}} + 2B + N_{\mathrm{in}} + 1\bigr)2^{16},
\qquad
Q_{\mathrm{dec}}=0,
\]
and deterministic local work equal to one execution of the fixed representative-table builds for stages $2,\dots,8$ plus, for each tested IV, a constant number of exact $2^{16}$-table passes together with the stage-dependent cycle, rotation, verifier, and stage-1 scans listed above.

As an empirical orientation point rather than an asymptotic claim, on the user's machine one measured full run with
\[
N_{\mathrm{out}}=10000,\qquad B=16,\qquad N_{\mathrm{in}}=1
\]
and the successful distinguished inward IV supplied on the command line completed in approximately $125.21$ seconds of wall-clock time.

\section{Implementation and runtime}
The attack is implemented in the multithreaded C program
\[
\texttt{separ\_public\_full\_hybrid\_attack.exe}.
\]
The public end-to-end runner is
\begin{verbatim}
separ_public_full_hybrid_attack.exe --search-trials N --search-beam B
    --search-seed S --inward-trials M --iv HEX32 --oracle-key HEX64
\end{verbatim}
with optional \texttt{--verbose} output. In the default mode the program shows only the current phase, progress, best IV, current IV, and best recovered key/state values in a compact in-place display.

Operationally, the cleaned public path:
\begin{itemize}[leftmargin=1.5em]
\item recovers $K_8$ only by the public hybrid score over the beam;
\item recovers stages $7$ through $2$ only by the exact public recursion above;
\item recovers stage~1 only by the exact closure of Theorem~\ref{thm:stage1};
\item accepts a full key only after exact equality of the recovered codebook with the oracle codebook.
\end{itemize}

\section{Conclusion}
The \texttt{main.c} implementation of SEPAR is broken end-to-end in the chosen-IV, chosen-plaintext, chosen-ciphertext, reset model. The critical structural fact is that, once a suitable IV is found, the implementation exposes a chain of exact public reduced tables in which each stage appears as a translated 16-bit permutation, and those translations can be stripped or solved exactly.

The resulting attack is not a heuristic key-ranking experiment. It is a deterministic full-key recovery pipeline conditioned on the success of the IV-search phase. The only remaining uncertainty is empirical: the probability that a random chosen IV falls into the recursive classes required by the outer bootstrap and the inward recursion.

\end{document}
