\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs}
\usepackage{array}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{verbatim}

\title{A Public Stagewise End-to-End Full-Key Recovery Attack on the \texttt{main.c} Implementation of SEPAR}
\author{}
\date{2026-04-15}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}
\newtheorem{proposition}{Proposition}

\newcommand{\Z}[1]{\mathbb{Z}_{#1}}
\newcommand{\hi}{\operatorname{hi}}
\newcommand{\lo}{\operatorname{lo}}
\newcommand{\rot}{\operatorname{rotl}_{16}}
\newcommand{\proj}{\operatorname{proj}}

\begin{document}
\maketitle

\begin{abstract}
We study the concrete SEPAR implementation in \texttt{SEPAR/main.c}, not an abstract idealized design. In the chosen-IV, chosen-plaintext, chosen-ciphertext, reset model, we give a public end-to-end full-key recovery attack on the 256-bit implementation key. The attack has a stochastic front-end and a deterministic exact core. The stochastic component is a chosen-IV search for suitable stage-8 and inward contexts. Once those IV sets are hit, every subsequent step is exact: an exact public stage-8 bootstrap for $K_8$; an exact stagewise inward search in which already-solved outer stages are peeled on fresh IVs by exact fixed-pair state recovery; and an exact stage-1 closure from the public stage-2 reduced table. Final acceptance is exact: a 256-bit key is accepted if and only if it reproduces the full $2^{16}$-entry one-word oracle table on the successful IV. We define the exact attack primitives used by the implementation, prove conditional exactness, report complete recovery on difficult random keys, and give a full-program Monte Carlo over 100 pseudorandom 256-bit keys. Under the explicit Monte Carlo outer-search profile $(N_{\mathrm{out}},B,\texttt{search-seed})=(512,32,1)$, the observed random-key success rates were $9/100$ at inward budget $N=2$, $45/100$ at $N=4$, $89/100$ at $N=8$, and $100/100$ at every tested budget $N\ge 16$.
\end{abstract}

\section{Introduction}
The target is the concrete word-oriented implementation in \texttt{SEPAR/main.c}. The relevant attack surface is the resettable 16-bit one-word oracle induced by that code. For every fixed IV, the implementation exposes a public permutation
\[
F_{\mathrm{IV}}:\Z{2^{16}}\to\Z{2^{16}}
\]
on one-word inputs. Because the block size is only 16 bits, the entire codebook is observable with $2^{16}$ chosen encryptions and can then be manipulated exactly.

The attack model is:
\begin{itemize}[leftmargin=1.5em]
\item chosen IV;
\item chosen plaintext and chosen ciphertext;
\item exact state reset between oracle calls.
\end{itemize}

Under that model, the implementation gives exact public access to
\[
F_{\mathrm{IV}}(x)=\operatorname{Enc}_{K,\mathrm{IV}}(x),
\qquad
F_{\mathrm{IV}}^{-1}(y)=\operatorname{Dec}_{K,\mathrm{IV}}(y),
\]
and the implementation used in the attack constructs the inverse locally by inverting the public encryption table.

\paragraph{Main claim.}
The public attack implemented in
\[
\texttt{separ\_public\_stagewise\_attack.c}
\]
recovers the full 256-bit key by:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item a chosen-IV beam search and exact hybrid public bootstrap for $K_8$;
\item an exact stagewise inward search for $K_7,K_6,\dots,K_2$ in which solved outer stages are peeled on fresh IVs by exact fixed-pair state recovery;
\item an exact stage-1 closure from the public stage-2 reduced table;
\item exact full-codebook verification of the recovered 256-bit key.
\end{enumerate}

What is proved below is \emph{conditional exactness}: once the sampled IV sets contain the true exact candidates required by the attack, the remainder of the recovery is deterministic and exact.

\section{Exact implementation model}
\subsection{Stage-local permutation}
Let the eight 32-bit stage keys be
\[
K_i=(k_{i,0},k_{i,1})\in \Z{2^{16}}\times \Z{2^{16}},
\qquad 1\le i\le 8.
\]

Let $\Sigma:\Z{2^{16}}\to\Z{2^{16}}$ be the nibblewise S-box layer induced by the fixed tables $S_1,S_2,S_3,S_4$ in \texttt{main.c}, and $\Sigma^{-1}$ the inverse layer induced by \texttt{IS1}, \dots, \texttt{IS4}.

Let $L:\Z{2^{16}}\to\Z{2^{16}}$ be the exact linear layer implemented by \texttt{sep\_rotl16}. If
\[
x=(a,b,c,d)\in(\Z{16})^4
\]
in nibble notation, define
\[
M(a,b,c,d)=(a\oplus c,\;b\oplus d,\;c\oplus b,\;d\oplus a),
\]
identify nibble tuples with 16-bit words, and set
\[
L(x)=M(x)\oplus \rot(M(x),12)\oplus \rot(M(x),8).
\]
Its inverse is exactly \texttt{sep\_inrotl16}.

For each stage $i$, the code derives two further 16-bit subkeys $\kappa_{i,2},\kappa_{i,3}$ from $(k_{i,0},k_{i,1})$:
\begin{align*}
u_i &= \rot(k_{i,0},6), &
v_i &= \rot(k_{i,1},10),\\
\kappa_{i,2} &= \bigl(u_i \vee (S_1((u_i \gg 6)\bmod 16)\ll 6)\bigr)\oplus (i+2),\\
\kappa_{i,3} &= \bigl(v_i \vee (S_1((v_i \gg 6)\bmod 16)\ll 6)\bigr)\oplus (i+3).
\end{align*}

With $A_t(x)=x\oplus t$, the stage-local encryption permutation implemented by \texttt{enc\_block} is exactly
\begin{equation}
\label{eq:stage-enc}
E_i
=
A_{\kappa_{i,2}\oplus \kappa_{i,3}}
\circ \Sigma
\circ A_{k_{i,0}\oplus k_{i,1}}
\circ L
\circ \Sigma
\circ A_{\kappa_{i,3}}
\circ L
\circ \Sigma
\circ A_{\kappa_{i,2}}
\circ L
\circ \Sigma
\circ A_{k_{i,1}}
\circ L
\circ \Sigma
\circ A_{k_{i,0}},
\end{equation}
and $D_i=E_i^{-1}$ is exactly \texttt{dec\_block}.

\subsection{Translated stage action and one-word encryption}
For a state word $s_i\in\Z{2^{16}}$, define the translated stage permutation
\[
B_i^{(s_i)}(x)=E_i(x+s_i),
\qquad
\bigl(B_i^{(s_i)}\bigr)^{-1}(y)=D_i(y)-s_i,
\]
where all additions and subtractions are in $\Z{2^{16}}$.

For a reset IV with state $(s_1,\dots,s_8)$, the public one-word encryption permutation is
\[
F_{\mathrm{IV}}
=
B_8^{(s_8)}
\circ B_7^{(s_7)}
\circ \cdots \circ
B_1^{(s_1)}.
\]

\subsection{State update law}
Write the eight state words as $(s_1,\dots,s_8)$ and the intermediate stage outputs as
\[
v_{12},v_{23},v_{34},v_{45},v_{56},v_{67},v_{78},c.
\]
From \texttt{Separ\_Encryption} in \texttt{main.c}, the exact state update after one word is
\begin{align}
s_2' &= s_2 + v_{12} + v_{56} + s_6, \label{eq:s2}\\
s_3' &= s_3 + v_{23} + v_{34} + s_4 + s_1, \label{eq:s3}\\
s_4' &= s_4 + v_{12} + v_{45} + s_8, \label{eq:s4}\\
s_5' &= s_5 + v_{23} + L', \label{eq:s5}\\
s_6' &= s_6 + v_{12} + v_{45} + s_7, \label{eq:s6}\\
s_7' &= s_7 + v_{23} + v_{67}, \label{eq:s7}\\
s_8' &= s_8 + v_{45}, \label{eq:s8}\\
s_1' &= s_1 + s_5 + v_{34} + 2v_{23} + v_{78}. \label{eq:s1}
\end{align}
The attack below uses only reset one-word tables, so these equations matter only because they determine which state word appears as the output translation at each stage.

\section{Exact public tables used by the attack}
\subsection{Recursive current tables}
For $m\in\{1,\dots,7\}$ define the inner cascade
\[
R_m
:=
B_m^{(s_m)}
\circ B_{m-1}^{(s_{m-1})}
\circ \cdots \circ
B_1^{(s_1)}.
\]

\begin{definition}[Current table at level $m$]
Assume $K_{m+1},\dots,K_8$ and $s_{m+2},\dots,s_8$ are known for a fixed IV. The current table at level $m$ is the public table
\[
C_m:\Z{2^{16}}\to\Z{2^{16}}
\]
computed recursively by
\[
C_7=D_8\circ F_{\mathrm{IV}},
\qquad
C_m=D_{m+1}\bigl(C_{m+1}-s_{m+2}\bigr)
\quad (1\le m\le 6).
\]
\end{definition}

\begin{lemma}[Exact translation form]
\label{lem:current-table}
For every $m\in\{1,\dots,7\}$,
\[
C_m(x)=R_m(x)+s_{m+1}
\qquad\text{for all }x\in\Z{2^{16}}.
\]
\end{lemma}

\begin{proof}
For $m=7$,
\[
F_{\mathrm{IV}}(x)=E_8(R_7(x)+s_8),
\]
hence
\[
C_7(x)=D_8(F_{\mathrm{IV}}(x))=R_7(x)+s_8.
\]
Assume the statement holds at level $m+1$. Then
\[
C_{m+1}(x)=R_{m+1}(x)+s_{m+2}
=
E_{m+1}(R_m(x)+s_{m+1})+s_{m+2},
\]
so
\[
C_m(x)
=
D_{m+1}(C_{m+1}(x)-s_{m+2})
=
R_m(x)+s_{m+1}.
\]
\end{proof}

Thus the exact public object attacked at level $m$ is the inner cascade $R_m$ observed through a single output translation by $s_{m+1}$.

\subsection{Exact low-byte score}
Let
\[
H(y)=\left\lfloor\frac{y}{2^8}\right\rfloor\in\Z{256}
\]
be the output high byte. For a public table $T:\Z{2^{16}}\to\Z{2^{16}}$, a candidate low byte $b\in\Z{256}$, and a row index $r\in\Z{256}$, define
\[
\sigma_{T,b}(r)
:=
\left|
\left\{
H\bigl(T(256r+\lambda)-b\bigr):\lambda\in\Z{256}
\right\}
\right|,
\]
and the exact total score
\[
\Sigma(T;b):=\sum_{r=0}^{255}\sigma_{T,b}(r).
\]
This is exactly the quantity computed by \texttt{exact\_low\_byte\_scan\_u16()}.

\subsection{Projected quotient cycles, representatives, and verifier}
Define the 4-bit projection
\[
\proj(y):=H(y)\bmod 16\in\Z{16}.
\]
For a stage key pair $K_i$, define the stage-$i$ quotient map
\[
q_{i,K_i}(h)=\proj(E_i(256h)),
\qquad h\in\Z{16},
\]
and the full high-byte quotient map
\[
Q_{i,K_i}(h)=H(E_i(256h)),
\qquad h\in\Z{256}.
\]

The implementation uses the exact auxiliary sets
\[
\Delta_0=\{1,2,4,8,15,16\},
\]
\[
\Omega=\{0,1,2,3,5,8,13,21,34,55,89,144,200,233,255\},
\]
\[
\mathcal R=\{0x00,0x55,0xAA,0xFF\}.
\]

The corrected implementation uses the same universal invisible-bit set for every stage $i\in\{2,\dots,8\}$:
\[
P_i=\{6,7,12,13,14,15,18,19,20,21,22,23\}.
\]
Hence
\[
|P_i|=12,\qquad |V_i|=20,
\]
and every representative table enumerates exactly
\[
2^{20}=1{,}048{,}576
\]
visible masks. Stages $2,\dots,7$ canonicalize 16-cycles by rotation to first entry $0$; stage 8 uses raw cycles.

For any public table $T$, define the exact projected edge-count matrix
\[
M_T(a,b)
:=
\left|
\left\{
(x,\delta)\in \Z{2^{16}}\times \Delta_0 :
\proj(T(x))=a,\;
\proj(T(x+\delta))=b
\right\}
\right|,
\qquad a,b\in\Z{16}.
\]
If $c=(c_0,\dots,c_{15})$ is a 16-cycle, define the exact cycle score
\[
S_T(c)=\sum_{j=0}^{15}M_T(c_j,c_{j+1\bmod 16}).
\]

To lift a projected cycle to a 256-point high-byte order, the implementation builds the exact 8-bit transition matrix
\[
N_T(u,v)
:=
\left|
\left\{
(x,\delta)\in \Z{2^{16}}\times \Delta_0 :
H(T(x))=u,\;
H(T(x+\delta))=v
\right\}
\right|,
\qquad u,v\in\Z{256},
\]
solves exact $16\times 16$ assignments between adjacent cycle layers by the Hungarian algorithm, and accepts only wrap permutations forming one 16-cycle.

Given a recovered cycle/order pair, the implementation enumerates every compatible representative-group base key and expands all invisible masks. It then checks exact equality against the reconstructed order targets first on $\Omega$ and then on all 256 values. Finally it exact-scores the surviving candidates by
\[
V_i(K_i;T)
:=
\sum_{r\in\mathcal R}
\min_{b\in\Z{256}}
\left|
\left\{
H\bigl(D_i(T(256r+\lambda))-b\bigr):\lambda\in\Z{256}
\right\}
\right|.
\]
This is the verifier used both in the outer bootstrap and in the stagewise inward search.

\section{The outer public bootstrap for \texorpdfstring{$K_8$}{K8}}
\subsection{Weak-IV beam search}
For an IV, define the exact row score
\[
\rho(\mathrm{IV})
:=
\left|
\left\{
H(F_{\mathrm{IV}}(\lambda)):\lambda\in\Z{256}
\right\}
\right|.
\]
The implementation keeps the $B$ IVs with smallest $\rho(\mathrm{IV})$. This is a search-ordering device only; it is not part of the acceptance criterion.

\subsection{Hybrid exact stage-8 recovery}
For each IV in the beam, the implementation:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item computes the exact stage-8 quotient/bootstrap candidate set by projected-cycle maximization and exact verifier scoring;
\item closes that set under the independent MSB toggles
\[
(a,b)\mapsto(a\oplus \epsilon_0 2^{15},\;b\oplus \epsilon_1 2^{15});
\]
\item exact-scores the resulting small pool by the additive differential score
\[
A_{\mathrm{IV}}(K_8)
:=
\sum_{\delta\in\Delta_0}
\max_{\Delta\in\Z{2^{16}}}
\left|
\left\{
x:
D_8(F_{\mathrm{IV}}(x+\delta);K_8)-D_8(F_{\mathrm{IV}}(x);K_8)=\Delta
\right\}
\right|;
\]
\item aggregates those exact scores across the beam and outputs the best $K_8$.
\end{enumerate}

\section{Exact stagewise inward search from \texorpdfstring{$K_7$}{K7} to \texorpdfstring{$K_2$}{K2}}
\subsection{Exact fixed-pair peeling on fresh IVs}
The inward phase is not a one-IV cascade. Once the outer bootstrap has fixed $K_8$, the program samples a fresh IV set independently at each target stage. On each such IV it peels every already-solved outer stage by exact state recovery for the fixed known stage key, and only then attacks the next unknown stage.

Fix a target stage $t\in\{7,6,5,4,3,2\}$ and suppose a branch has already fixed
\[
K_{t+1},K_{t+2},\dots,K_8.
\]
Let $\eta$ be a fresh IV. Define
\[
C_7^{(\eta)}:=D_8\circ F_\eta.
\]
If $K_8$ is correct, then by Lemma~\ref{lem:current-table},
\[
C_7^{(\eta)}(x)=R_7^{(\eta)}(x)+s_8^{(\eta)}.
\]

For a fixed stage $m\in\{7,\dots,t+1\}$ and a public table $T$, define the minimizing low-byte set
\[
L_m(T):=\arg\min_{b\in\Z{256}}\Sigma(T;b).
\]
For a fixed known pair $K_m$, define the exact fixed-pair state solver
\begin{equation}
\label{eq:fixed-pair-state-solver}
\mathcal{S}_m(T;K_m)
:=
\arg\min_{\substack{s=b+256h\\ b\in L_m(T),\,h\in\Z{256}}}
V_m(K_m;T-s),
\end{equation}
with lexicographic tie-break on $s$. This is exactly \texttt{recover\_known\_pair\_state\_fast()}.

Starting from $C_7^{(\eta)}$, the program peels already-solved outer stages by
\begin{align}
\widehat{s}_{m+1}^{(\eta)} &:= \mathcal{S}_m(C_m^{(\eta)};K_m), \label{eq:stagewise-state}\\
C_{m-1}^{(\eta)} &:= D_m\!\left(C_m^{(\eta)}-\widehat{s}_{m+1}^{(\eta)};K_m\right), \qquad m=7,6,\dots,t+1. \label{eq:stagewise-peel}
\end{align}
If every known outer pair on the current branch is correct and every fixed-pair state recovery returns the true state word, then
\[
C_t^{(\eta)}(x)=R_t^{(\eta)}(x)+s_{t+1}^{(\eta)}.
\]

\subsection{Exact target-stage candidate sets}
For a fixed target stage $t$ and fresh IV $\eta$, let
\[
\mathcal{A}_t(\eta;K_{t+1},\dots,K_8)
\]
denote the set of all minimum-verifier rows produced by applying the exact stage-$t$ scan of Sections~3.2--3.4 to the prepared table $C_t^{(\eta)}$. Every element of $\mathcal{A}_t$ is a pair
\[
(K_t,s_{t+1})\in(\Z{2^{16}})^2.
\]
Write
\[
\Pi_{\mathrm{pair}}\mathcal{A}_t(\eta;K_{t+1},\dots,K_8)
\]
for the set of stage-key pairs appearing in those rows.

Let
\[
\Omega_t(N)=\{\eta_{t,1},\dots,\eta_{t,N}\}
\]
be the first $N$ fresh IVs in the deterministic stage-$t$ IV stream. Define the exact stagewise candidate-pair set
\begin{equation}
\label{eq:stagewise-pair-set}
\mathcal{P}_t^{(N)}(K_{t+1},\dots,K_8)
:=
\bigcup_{\eta\in\Omega_t(N)}
\Pi_{\mathrm{pair}}\mathcal{A}_t(\eta;K_{t+1},\dots,K_8).
\end{equation}
The implementation also records the exact hit multiplicity
\begin{equation}
\label{eq:stagewise-hit-count}
h_t^{(N)}(K_t)
:=
\left|
\left\{
\eta\in\Omega_t(N):
K_t\in\Pi_{\mathrm{pair}}\mathcal{A}_t(\eta;K_{t+1},\dots,K_8)
\right\}
\right|,
\end{equation}
sorts candidates by decreasing $h_t^{(N)}$, and recursively explores all distinct pairs in that exact order. This ordering affects runtime only: no candidate in~\eqref{eq:stagewise-pair-set} is discarded.

\subsection{Exact recursive search tree}
Define recursively the exact search procedures
\[
\mathfrak{S}_t^{(N)}(K_{t+1},\dots,K_8),
\qquad t=7,6,\dots,2,
\]
as follows.

For $t\in\{7,6,5,4,3\}$:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item build $\mathcal{P}_t^{(N)}(K_{t+1},\dots,K_8)$ by~\eqref{eq:stagewise-pair-set};
\item enumerate all $K_t\in\mathcal{P}_t^{(N)}$ in decreasing order of~\eqref{eq:stagewise-hit-count};
\item recurse to
\[
\mathfrak{S}_{t-1}^{(N)}(K_t,K_{t+1},\dots,K_8).
\]
\end{enumerate}

For $t=2$, there is no further pair recursion. Instead, for every $\eta\in\Omega_2(N)$ the program:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item prepares the exact public stage-2 current table $C_2^{(\eta)}$ by fixed-pair peeling of the already-solved suffix;
\item computes the exact minimum-verifier row set $\mathcal{A}_2(\eta;K_3,\dots,K_8)$;
\item for every row $(K_2,s_3)\in \mathcal{A}_2(\eta;K_3,\dots,K_8)$, invokes the exact stage-1 closure of Section~6;
\item accepts only if the recovered 256-bit key reproduces the entire $2^{16}$-entry one-word oracle table on the same IV.
\end{enumerate}

\paragraph{Why stage 1 is separate.}
The projected-family scan exists only for stages $2,\dots,8$. Stage 1 has no visible successor stage and therefore no stage-1 quotient-cycle analogue. It is closed exactly from the stage-2 reduced table.

\begin{proposition}[Monotonicity in the inward budget]
\label{prop:monotone-budget}
Fix a key, the outerly recovered $K_8$, and the deterministic per-stage IV streams. If the stagewise attack succeeds with inward budget $N$, then it also succeeds with every larger budget $N'\ge N$.
\end{proposition}

\begin{proof}
For every stage $t$, the sampled IV set $\Omega_t(N)$ is a prefix of $\Omega_t(N')$. Hence
\[
\mathcal{P}_t^{(N)}(K_{t+1},\dots,K_8)
\subseteq
\mathcal{P}_t^{(N')}(K_{t+1},\dots,K_8)
\]
for every fixed suffix. Thus every branch explored at budget $N$ is also explored at budget $N'$. At stage 2 the tested IV set is likewise prefix-monotone, so any successful stage-2 IV at budget $N$ is still tested at budget $N'$. Therefore success is monotone.
\end{proof}

\section{Exact stage-1 closure from the public stage-2 table}
\subsection{Reduction to a single translated stage}
Assume the stagewise recursion has recovered the true $(K_2,s_3)$ on some successful stage-2 IV. Then
\[
C_2(x)=E_2(E_1(x+s_1)+s_2)+s_3.
\]
Define
\[
W_2(x)=D_2(C_2(x)-s_3;K_2).
\]
Then exactly
\begin{equation}
\label{eq:stage1-reduction}
W_2(x)=E_1(x+s_1)+s_2.
\end{equation}

\subsection{Low-byte recovery for \texorpdfstring{$s_2$}{s2}}
The implementation first applies the same exact low-byte scan to $W_2$. For each candidate
\[
s_2^\star\in\Z{2^{16}},
\]
it forms
\[
C^\star(x)=W_2(x)-s_2^\star.
\]
Any true stage-1 solution must satisfy
\[
C^\star(x)=E_1(x+s_1;K_1).
\]

\subsection{Step 1: exact row-signature inversion}
For $u\in\Z{256}$ define the 16-point row signature
\[
R_u^{C^\star}(h)=\proj(C^\star(u+256h)),
\qquad h\in\Z{16}.
\]
Write
\[
s_1=s_{1,\mathrm{lo}}+256\,s_{1,\mathrm{mid}}+4096\,s_{1,\mathrm{hi}},
\]
with $s_{1,\mathrm{lo}}\in\Z{256}$ and $s_{1,\mathrm{mid}},s_{1,\mathrm{hi}}\in\Z{16}$.

For a candidate $s_{1,\mathrm{lo}}$, the routine \texttt{stage1\_step1\_candidates\_u16()} checks the exact carry partition:
\begin{quote}
all rows with $u+s_{1,\mathrm{lo}}<256$ must share one common signature $\sigma$, and all rows with $u+s_{1,\mathrm{lo}}\ge 256$ must share the one-step rotation $\operatorname{rot}_1(\sigma)$.
\end{quote}
For each surviving $s_{1,\mathrm{lo}}$ and each $s_{1,\mathrm{mid}}\in\Z{16}$, the implementation rotates $\sigma$ by $16-s_{1,\mathrm{mid}}$ and inverts it through the exact precomputed family
\[
\nu_{a,b,c,d}:\Z{16}\to\Z{16},
\]
where
\[
\nu_{a,b,c,d}(x)=S_2\bigl(t_4(x)\oplus a\oplus b\bigr)\oplus c\oplus d.
\]
with
\begin{align*}
t_1(x)&=S_2(x\oplus a),\\
t_2(x)&=S_2(t_1(x)\oplus b),\\
t_3(x)&=S_2(t_2(x)\oplus c),\\
t_4(x)&=S_2(t_3(x)\oplus d).
\end{align*}
This yields exact candidates for
\[
s_{1,\mathrm{lo}},\quad s_{1,\mathrm{mid}},\quad a,\quad b,\quad c,\quad d.
\]

\subsection{Step 2: exact lift to full \texorpdfstring{$K_1$}{K1} candidates}
The routine next uses exact inverse-image tables for the derived-key high nibbles of $\kappa_{1,2}$ and $\kappa_{1,3}$. For each surviving $(s_{1,\mathrm{lo}},s_{1,\mathrm{mid}},a,b,c,d)$, each
\[
\ell_6\in I_2(c),\qquad \beta\in I_3(d),
\]
and each 12-bit free mask $f\in\Z{2^{12}}$, it constructs
\begin{align}
k_{1,0}
&=
\ell_6
\;\vee\;
\bigl((f\bmod 4)\ll 6\bigr)
\;\vee\;
(a\ll 8)
\;\vee\;
\left(\left(\left\lfloor\frac{f}{4}\right\rfloor\bmod 16\right)\ll 12\right), \label{eq:k10-lift}\\
k_{1,1}
&=
(\beta\bmod 4)
\;\vee\;
\left(\left(\left\lfloor\frac{f}{64}\right\rfloor\bmod 64\right)\ll 2\right)
\;\vee\;
(b\ll 8)
\;\vee\;
\left(\left(\left\lfloor\frac{\beta}{4}\right\rfloor\bmod 16\right)\ll 12\right). \label{eq:k11-lift}
\end{align}

The resulting pair is exact-tested on the 12-bit slice
\[
X_{12}=\{0,\dots,4095\}
\]
by comparing nibble $(7..4)$ values:
\[
\left\lfloor\frac{E_1((x+s_{1,\mathrm{lo}}+256s_{1,\mathrm{mid}})\bmod 2^{12};K_1)}{2^4}\right\rfloor\bmod 16
=
\left\lfloor\frac{C^\star(x)}{2^4}\right\rfloor\bmod 16
\qquad\text{for all }x\in X_{12}.
\]

\subsection{Step 3: exact recovery of the top nibble of \texorpdfstring{$s_1$}{s1}}
For each surviving pair $K_1$, the implementation scans all 16 possibilities for $s_{1,\mathrm{hi}}$ and performs full exact verification:
\[
E_1(x+s_1;K_1)=C^\star(x)
\qquad\text{for all }x\in\Z{2^{16}}.
\]
If exactly one triple $(K_1,s_1,s_2)$ survives, the stage-1 closure succeeds.

\begin{theorem}[Exactness of the stage-1 closure]
\label{thm:stage1}
Let $W_2$ be the exact public table of~\eqref{eq:stage1-reduction}. If there is a unique triple $(K_1,s_1,s_2)$ satisfying
\[
W_2(x)=E_1(x+s_1;K_1)+s_2
\qquad\text{for all }x\in\Z{2^{16}},
\]
then \texttt{stage1\_finish\_from\_stage2\_current()} returns exactly that triple.
\end{theorem}

\begin{proof}
The routine is exhaustive over all candidates for the low and high bytes of $s_2$, all exact row-signature solutions for $(s_{1,\mathrm{lo}},s_{1,\mathrm{mid}},a,b,c,d)$, all inverse-image candidates from $I_2(c)$ and $I_3(d)$, all $2^{12}$ free masks in~\eqref{eq:k10-lift}--\eqref{eq:k11-lift}, and all 16 possibilities for $s_{1,\mathrm{hi}}$. Every surviving candidate is tested by exact equality either on the 12-bit slice or on the full $2^{16}$-entry table, and the routine accepts only if the number of full solutions is exactly one. Hence a unique solution is returned exactly.
\end{proof}

\section{Public full-key acceptance and conditional correctness}
\begin{definition}[Exact public acceptance test]
Given recovered pairs
\[
\widehat{K}=(\widehat{K}_1,\dots,\widehat{K}_8),
\]
rebuild the candidate one-word table
\[
\widehat{F}_{\mathrm{IV}}
\]
under the same IV and accept if and only if
\[
\widehat{F}_{\mathrm{IV}}=F_{\mathrm{IV}}
\]
as functions on $\Z{2^{16}}$.
\end{definition}

Because the block size is only 16 bits, this is an exact comparison of two $65536$-entry tables.

\begin{theorem}[Conditional exactness of the full public attack]
\label{thm:main}
Fix the exact beam IV set used by the outer stage-8 search and the exact fresh-IV sets $\Omega_t(N)$ used by the stagewise search. Assume:
\begin{enumerate}[label=(\alph*),leftmargin=1.5em]
\item the outer hybrid bootstrap ranks the true $K_8$ first;
\item for each $t\in\{7,6,5,4,3\}$, on the branch
\[
(K_{t+1},\dots,K_8)=(K_{t+1}^{\star},\dots,K_8^{\star}),
\]
the true pair $K_t^{\star}$ belongs to the exact candidate set
\[
\mathcal{P}_t^{(N)}(K_{t+1}^{\star},\dots,K_8^{\star});
\]
\item for some $\eta\in\Omega_2(N)$ on the true branch, the exact stage-2 scan contains the true row $(K_2^{\star},s_3^{\star})$, and the stage-1 system
\[
W_2(x)=E_1(x+s_1;K_1)+s_2
\]
has a unique solution $(K_1^{\star},s_1^{\star},s_2^{\star})$.
\end{enumerate}
Then the implementation outputs the true 256-bit key and accepts it by the exact codebook test. No false full key can be accepted.
\end{theorem}

\begin{proof}
By assumption~(a), the recovered outer pair is the true $K_8^{\star}$. Consider the recursive branch
\[
K_7^{\star},K_6^{\star},\dots,K_3^{\star}.
\]
At stage $t$, assumption~(b) says that
\[
K_t^{\star}\in \mathcal{P}_t^{(N)}(K_{t+1}^{\star},\dots,K_8^{\star}).
\]
Since the implementation eventually explores every candidate in that set, the true branch is never pruned. Therefore the recursive search reaches stage 2 on the true suffix.

Now fix the stage-2 IV $\eta$ from assumption~(c). On the true branch, the fixed-pair peeling equations~\eqref{eq:stagewise-state}--\eqref{eq:stagewise-peel} reconstruct the exact public current table
\[
C_2^{(\eta)}(x)=R_2^{(\eta)}(x)+s_3^{(\eta)}.
\]
Since the exact stage-2 scan includes the true row $(K_2^{\star},s_3^{\star})$, the reduction~\eqref{eq:stage1-reduction} holds for the true reduced table
\[
W_2(x)=D_2(C_2^{(\eta)}(x)-s_3^{\star};K_2^{\star}).
\]
By assumption~(c) and Theorem~\ref{thm:stage1}, the stage-1 closure returns exactly the unique true triple $(K_1^{\star},s_1^{\star},s_2^{\star})$. Hence the reconstructed full key is the true implementation key. Finally, the acceptance test compares complete $2^{16}$-entry tables, so no false full key can be accepted.
\end{proof}

\section{Worked random-key recovery}
\subsection{Target key and successful IVs}
As a representative difficult instance, consider the random key
\[
K=
\texttt{8FCD06ACB99FB6F77135F7ABCD27778AD08489BA2C6DE01CFC17FF0CE0F553C1}.
\]

The successful public stagewise run used
\begin{itemize}[leftmargin=1.5em]
\item outer beam budget $N_{\mathrm{out}}=512$;
\item beam size $B=32$;
\item search seed $1$;
\item inward budget $N=64$ per target stage;
\item successful stage-2 IV
\[
\texttt{F71831A507C6103EBB51EC0AC07A0C60}.
\]
\end{itemize}

\subsection{Recovered values}
The public recovery returned
\begin{align*}
K_8 &= (\texttt{E0F5},\texttt{53C1}),\\
K_7 &= (\texttt{FC17},\texttt{FF0C}),\qquad s_8=\texttt{675F},\\
K_6 &= (\texttt{2C6D},\texttt{E01C}),\qquad s_7=\texttt{E454},\\
K_5 &= (\texttt{D084},\texttt{89BA}),\qquad s_6=\texttt{328C},\\
K_4 &= (\texttt{CD27},\texttt{778A}),\qquad s_5=\texttt{8D1B},\\
K_3 &= (\texttt{7135},\texttt{F7AB}),\qquad s_4=\texttt{A2CF},\\
K_2 &= (\texttt{B99F},\texttt{B6F7}),\qquad s_3=\texttt{6DEA},\\
K_1 &= (\texttt{8FCD},\texttt{06AC}),\qquad s_2=\texttt{7F43},\qquad s_1=\texttt{27B3}.
\end{align*}

The decisive public log line is
\begin{verbatim}
full-key candidate verified against exact one-word oracle table
\end{verbatim}
from \texttt{stagewise\_attack\_hard1\_default64.log}.

\section{Monte Carlo estimate for full-program random-key success}
\subsection{Experiment design}
This Monte Carlo evaluates the entire implemented attack, not a conditional inward probe. We sampled
\[
m=100
\]
independent pseudorandom 256-bit keys by
\[
K_j:=\operatorname{SHA256}(\texttt{``stagewise-mc-key-}j\texttt{''}),
\qquad j=0,\dots,99.
\]
To isolate the dependence on the inward budget, we fixed the outer search profile to
\[
N_{\mathrm{out}}=512,\qquad B=32,\qquad \texttt{search-seed}=1,\qquad \texttt{workers}=16,
\]
independently of the current binary defaults. We then evaluated the full public program
\[
\texttt{separ\_public\_stagewise\_attack.exe}
\]
at inward budgets
\[
N\in\{2,4,8,16,32,64\}.
\]
The budget-$64$ sample was the already completed 100-key run logged in
\[
\texttt{monte\_carlo\_stagewise\_100.log}.
\]
For budgets $2,4,8,16,32$ we reran the same 100-key set and copied the existing budget-$64$ rows verbatim into the combined dump
\[
\texttt{monte\_carlo\_stagewise\_budget\_sweep.log}.
\]
Because of Proposition~\ref{prop:monotone-budget}, success is monotone in $N$. Accordingly, within the budget sweep, once a key succeeded at a smaller budget, all larger budgets were recorded as monotone skips rather than rerun.

\subsection{Observed success rates}
Let
\[
P_N:=\Pr[\text{full public attack succeeds on a random key at inward budget }N]
\]
under the fixed outer parameters
\[
(N_{\mathrm{out}},B,\texttt{search-seed})=(512,32,1).
\]
For each budget we report the observed success rate together with the exact two-sided Clopper--Pearson 95\% confidence interval. The measured success table is:

\begin{center}
\small
\begin{tabular}{>{\raggedleft\arraybackslash}p{1.3cm}>{\raggedleft\arraybackslash}p{2.8cm}>{\raggedleft\arraybackslash}p{2.8cm}>{\raggedleft\arraybackslash}p{1.5cm}>{\raggedleft\arraybackslash}p{2.5cm}}
\toprule
Budget $N$ per stage & Total IV contexts & Chosen encryptions & Successes & 95\% exact CI \\
\midrule
2   & $512+32+6\cdot 2=556$   & $36{,}438{,}016$ & $9/100$   & $[0.0420,\,0.1640]$ \\
4   & $512+32+6\cdot 4=568$   & $37{,}224{,}448$ & $45/100$  & $[0.3503,\,0.5527]$ \\
8   & $512+32+6\cdot 8=592$   & $38{,}797{,}312$ & $89/100$  & $[0.8117,\,0.9438]$ \\
16  & $512+32+6\cdot 16=640$  & $41{,}943{,}040$ & $100/100$ & $[0.9638,\,1]$ \\
32  & $512+32+6\cdot 32=736$  & $48{,}234{,}496$ & $100/100$ & $[0.9638,\,1]$ \\
64  & $512+32+6\cdot 64=928$  & $60{,}817{,}408$ & $100/100$ & $[0.9638,\,1]$ \\
\bottomrule
\end{tabular}
\end{center}

Thus the random-key success curve rises sharply with the inward budget and reaches empirical saturation by $N=16$ on the 100-key sample. At every tested budget $N\ge 16$, the observed success rate is $100\%$, with exact 95\% lower confidence bound
\[
\underline{P}_{95}(100/100)=0.963783307\ldots
\]
and corresponding one-sided statistical error bar
\[
1-0.963783307\approx 0.036216693,
\]
that is, about $3.62$ percentage points.

\subsection{Runtime on the Monte Carlo sample}
For the independently executed budget-$64$ sample of 100 keys, the measured wall-clock times on the user's machine were:
\begin{center}
\begin{tabular}{rr}
\toprule
Statistic & Seconds \\
\midrule
mean   & $52.520$ \\
median & $52.346$ \\
min    & $30.358$ \\
max    & $89.096$ \\
\bottomrule
\end{tabular}
\end{center}
Thus, under the explicit Monte Carlo profile $(N_{\mathrm{out}},B,N)=(512,32,64)$, the full public attack is a seconds-to-minutes computation rather than a brute-force search.

\section{Complexity and bound}
\subsection{Oracle-query complexity}
One full one-word codebook requires exactly
\[
2^{16}
\]
chosen encryptions. The implementation does not require chosen-ciphertext queries, so
\[
Q_{\mathrm{dec}}=0.
\]

For outer budget $N_{\mathrm{out}}$, beam size $B$, and per-stage inward budget $N$, the successful direct path of the stagewise attack builds:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item $N_{\mathrm{out}}$ outer-search codebooks;
\item $B$ exact beam codebooks for the hybrid $K_8$ bootstrap;
\item $N$ fresh-IV codebooks for each target stage $t=7,6,5,4,3,2$.
\end{enumerate}
Hence the exact chosen-encryption count on a no-backtracking successful path is
\begin{equation}
\label{eq:qenc-direct}
Q_{\mathrm{enc}}^{\mathrm{direct}}
=
\bigl(N_{\mathrm{out}}+B+6N\bigr)2^{16}.
\end{equation}
With the current binary defaults
\[
N_{\mathrm{out}}=64,\qquad B=8,\qquad N=64,
\]
this becomes
\[
Q_{\mathrm{enc}}^{\mathrm{direct}}
=(64+8+384)2^{16}
=456\cdot 65536
=29{,}884{,}416.
\]

\subsection{Exact branch-aware bound}
The implementation is recursive and may, in principle, backtrack over exact candidate-pair sets before reaching the accepting branch. To state the exact implemented bound, let
\[
M_t(N)
\]
be the maximum, over all keys and all previously fixed suffixes $(K_{t+1},\dots,K_8)$ reachable by the program, of the cardinality of the exact stagewise candidate set
\[
\mathcal{P}_t^{(N)}(K_{t+1},\dots,K_8).
\]
Define recursively the maximum number of fresh-IV codebooks visited below stage $t$ by
\begin{align}
T_2(N) &:= N, \label{eq:T2}\\
T_t(N) &:= N + M_t(N)\,T_{t-1}(N), \qquad t=3,4,5,6,7. \label{eq:Trec}
\end{align}
Then the exact program-level chosen-encryption bound is
\begin{equation}
\label{eq:qenc-branch}
Q_{\mathrm{enc}}
=
\bigl(N_{\mathrm{out}}+B+T_7(N)\bigr)2^{16},
\qquad
Q_{\mathrm{dec}}=0.
\end{equation}
Equation~\eqref{eq:qenc-branch} is the correct exact bound for the implemented recursive search tree. Equation~\eqref{eq:qenc-direct} is the realized cost when the true branch is reached without backtracking. In the 100-key Monte Carlo sweep, this no-backtracking behavior was observed on every successful key.

\subsection{Exact local work}
Once a codebook is built, all remaining work is local and exact. For each stage-$t$ fresh IV, the local work consists of:
\begin{enumerate}[label=\arabic*.,leftmargin=1.5em]
\item exact fixed-pair state recovery for each already-solved stage above $t$, using~\eqref{eq:fixed-pair-state-solver};
\item one exact stage-$t$ low-byte scan;
\item one exact maximizing projected-cycle search over the $2^{20}$-entry representative space;
\item exact order reconstruction by Hungarian assignment lifts;
\item exact candidate lifting and exact verifier minimization;
\item for $t=2$, the exact stage-1 closure of Section~6 together with full-table verification.
\end{enumerate}

\subsection{Overall bound}
For fixed budgets $(N_{\mathrm{out}},B,N)$, the exact implemented query bound is~\eqref{eq:qenc-branch}. On the empirically dominant no-backtracking path, it collapses to the direct-path cost~\eqref{eq:qenc-direct}. For the default parameters of the current binary, the direct chosen-encryption count is
\[
29{,}884{,}416.
\]
Under the explicit Monte Carlo profile $(N_{\mathrm{out}},B,N)=(512,32,64)$, the 100-key sample averaged $52.520$ seconds per full key recovery, with median $52.346$ seconds and worst case $89.096$ seconds.

\section{Implementation and runtime}
The attack is implemented in the multithreaded C program
\[
\texttt{separ\_public\_stagewise\_attack.exe}.
\]
The public end-to-end runner is
\begin{verbatim}
separ_public_stagewise_attack.exe --search-trials N --search-beam B
    --search-seed S --inward-trials M --workers W --oracle-key HEX64
\end{verbatim}
with optional \texttt{--verbose} output. In the default mode the program shows only the current phase, progress, best IV, and recovered key segments in a compact in-place display.

Operationally, the public path:
\begin{itemize}[leftmargin=1.5em]
\item recovers $K_8$ only by the public hybrid score over the beam;
\item recovers stages $7$ through $2$ only by the exact stagewise recursion of Section~5;
\item recovers stage 1 only by the exact closure of Theorem~\ref{thm:stage1};
\item accepts a full key only after exact equality of the recovered codebook with the oracle codebook.
\end{itemize}

The default parameters of the current binary are
\[
N_{\mathrm{out}}=64,\qquad B=8,\qquad N=64,\qquad W=16.
\]
The full-program Monte Carlo of Section~8 used the explicit larger outer-search profile
\[
N_{\mathrm{out}}=512,\qquad B=32,\qquad \texttt{search-seed}=1,\qquad W=16
\]
in order to measure the dependence of success probability on the inward budget $N$ alone.

\section{Conclusion}
The \texttt{main.c} implementation of SEPAR is broken end-to-end in the chosen-IV, chosen-plaintext, chosen-ciphertext, reset model. The decisive structural fact is that every stage is exposed through an exact translated $2^{16}$-point permutation, and once a suffix stage key is fixed the corresponding IV-local state word can itself be recovered exactly and peeled away on fresh IVs. This yields the stagewise recursion of Section~5 and removes the dependence on one exceptionally good inward IV.

The attack is therefore not a heuristic key-ranking experiment. Conditional on the sampled IV sets containing the true exact candidates, the recovery is deterministic and exact; false full keys are excluded by complete one-word table equality. On a 100-key Monte Carlo sample of the full program, the explicit profile $(N_{\mathrm{out}},B,\texttt{search-seed})=(512,32,1)$ achieved success rates $0.09,0.45,0.89,1,1,1$ at inward budgets $N=2,4,8,16,32,64$, respectively, with exact 95\% lower confidence bound $0.9638$ once $N\ge 16$.


\end{document}
