\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{enumitem}

\title{An Exact Right-Recursive Ladder Attack on the Broken SEPAR Implementation}
\author{}
\date{}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

\begin{document}
\maketitle

\begin{abstract}
We study the \emph{implementation} of SEPAR in \texttt{main.c}, not the paper design. The critical implementation error is that the purported 16-bit diffusion layer collapses exactly to
\[
L(a,b,c,d)=(b\oplus c,\; b,\; a,\; a\oplus d),
\]
which makes every stage byte-triangular and leaves an autonomous 4-bit lane inside every \texttt{ENC\_Block}. Combined with the matched-context forward/inverse oracle exposed by the word-oriented mode, this yields a Hummingbird-style divide-and-conquer attack. The main contribution of this note is an exact \emph{right-recursive ladder}: after peeling the already known outer right stages, one can inject a candidate block permutation for the next visible stage directly into the reduced inverse oracle. For the true stage key, the current state word cancels from the induced additive differential spectrum exactly. In the state-injected model this exact ladder recovers all segment keys $K_8,\dots,K_1$. We further show that the 16-bit LFSR contributes no secrecy to same-reset differential attacks and that, once the four-word outer-even core $(s_2,s_4,s_6,s_8)$ is known at two consecutive rounds, the remaining state words $(s_1,s_3,s_5,s_7)$ and the stepped LFSR word are recovered by explicit algebraic formulas.
\end{abstract}

\section{Introduction}
The implementation of SEPAR is a cascade of eight 16-bit keyed permutations:
\[
F = B_8 \circ B_7 \circ \cdots \circ B_1,
\qquad
B_i(x)=E_i(x+s_i),
\]
where $E_i$ is the stage-local implementation of \texttt{ENC\_Block}, $s_i\in \mathbb{Z}_{2^{16}}$ is the current state word for stage $i$, and all additions/subtractions are modulo $2^{16}$.

Earlier work on this code base established three structural weaknesses.
\begin{enumerate}[label=(\roman*)]
\item \textbf{Exact byte triangularity.} The upper output byte of $E_i$ depends only on the upper input byte.
\item \textbf{Autonomous nibble lane.} The second output nibble of $E_i$ depends only on the second input nibble, giving a true 4-bit keyed permutation channel.
\item \textbf{Matched-context oracle.} For any chosen prefix, encryption and decryption expose a matched 16-bit permutation pair $(F,F^{-1})$ for the next word.
\end{enumerate}

The remaining question was whether these primitives can be organized into a genuinely \emph{exact} recursive attack in the style of the classical Hummingbird-1 break, rather than a beam-search engineering attack. This note answers that question affirmatively on the right side of the cascade.

\section{The implementation bug}
Write a 16-bit word in nibble form as $(a,b,c,d)$ from most significant to least significant nibble. The implementation of \texttt{Sep\_ROTL16} first performs
\[
a\leftarrow a\oplus c,\quad
b\leftarrow b\oplus d,\quad
c\leftarrow c\oplus b,\quad
d\leftarrow d\oplus a,
\]
and then xors the result with two rotations.

\begin{theorem}
The implemented linear layer satisfies
\[
L(a,b,c,d)=(b\oplus c,\; b,\; a,\; a\oplus d).
\]
\end{theorem}

\begin{proof}
This is a direct nibble-wise simplification of the code path in \texttt{Sep\_ROTL16}. The resulting expression is exact and was exhaustively verified in code.
\end{proof}

Two immediate consequences are central.
\begin{corollary}
For each stage $i$ there exists an 8-bit family of functions $U_i,L_i$ such that
\[
E_i(u\| \ell)=U_i(u)\,\|\,L_i(u,\ell),
\]
where $u$ is the upper input byte and $\ell$ is the lower input byte.
\end{corollary}

\begin{corollary}
For each stage $i$ there exists a 4-bit permutation $f_i$ such that the second output nibble of $E_i$ is exactly
\[
\operatorname{nib}_2(E_i(a,b,c,d))=f_i(b).
\]
\end{corollary}

Thus the full 16-bit block does \emph{not} behave like a well-diffused SP-network. It behaves like a triangular machine whose upper-half branching is driven almost entirely by external carries.

\section{Matched-context forward and inverse oracles}
Let $P=(p_0,\dots,p_{m-1})$ be a chosen plaintext prefix and let $C=(c_0,\dots,c_{m-1})$ be its ciphertext prefix under a fixed IV. Because the encryption and decryption code paths use the \emph{same} post-word state update, the internal state before the next word is the same in
\[
\text{Enc}(P\|x)
\qquad\text{and}\qquad
\text{Dec}(C\|y).
\]
Hence we obtain a matched permutation pair
\[
F_{P,\mathrm{IV}}(x)=\text{Enc}(P\|x)[m],
\qquad
G_{C,\mathrm{IV}}(y)=\text{Dec}(C\|y)[m]=F_{P,\mathrm{IV}}^{-1}(y).
\]

This is the exact 16-bit handle used throughout the attack. In practice one builds the full codebook of $F$ with $2^{16}$ chosen encryption queries and the full codebook of $G$ with the corresponding chosen decryption queries.

\section{Stage notation}
Fix a matched context and suppress the context subscripts. Write
\[
F = B_8\circ B_7\circ \cdots \circ B_1,
\qquad
B_i(x)=E_i(x+s_i).
\]
The inverse stage is
\[
B_i^{-1}(y)=D_i(y)-s_i,
\]
where $D_i=E_i^{-1}$ is the stage-local \texttt{DEC\_Block}.

For $m\in\{1,\dots,8\}$ define the right-reduced inverse oracle
\[
H_m = G\circ B_8\circ B_7\circ\cdots\circ B_{m+1}.
\]
When the already peeled outer right stages are correct, this telescopes to
\[
H_m = B_1^{-1}\circ B_2^{-1}\circ\cdots\circ B_m^{-1}.
\]

\section{The exact injected ladder}
The core exact identity of this note is obtained by injecting a candidate stage permutation directly into the reduced inverse oracle.

\begin{definition}[Injected stage test]
For a candidate key $K$ for stage $m$, define
\[
J_{m,K}(z)=H_m(E_m(z;K)).
\]
\end{definition}

\begin{theorem}[Exact state cancellation]
Let $K_m$ be the true segment key of stage $m$. Then
\[
J_{m,K_m}(z)=Q_{m-1}^{-1}(z-s_m),
\]
where
\[
Q_{m-1}=B_{m-1}\circ B_{m-2}\circ\cdots\circ B_1
\]
is the inner $(m-1)$-stage cascade.
\end{theorem}

\begin{proof}
Substitute $K=K_m$:
\[
J_{m,K_m}(z)=H_m(E_m(z;K_m)).
\]
Since
\[
H_m = B_1^{-1}\circ \cdots \circ B_m^{-1},
\qquad
B_m^{-1}(y)=D_m(y)-s_m,
\]
we obtain
\[
B_m^{-1}(E_m(z;K_m))=z-s_m.
\]
Therefore
\[
J_{m,K_m}(z)=B_1^{-1}\circ\cdots\circ B_{m-1}^{-1}(z-s_m)=Q_{m-1}^{-1}(z-s_m).
\qedhere
\]
\end{proof}

The decisive feature is that the current state word $s_m$ appears only as an \emph{input translation}. Hence additive differentials of $J_{m,K_m}$ are translation-invariant:
\[
J_{m,K_m}(z+\Delta)-J_{m,K_m}(z)
=
Q_{m-1}^{-1}(z+\Delta-s_m)-Q_{m-1}^{-1}(z-s_m).
\]
The shift by $s_m$ cancels.

\subsection{Exact stage score}
For a fixed difference set $\mathcal{D}\subset \mathbb{Z}_{2^{16}}^\ast$, define
\[
S_m(K)=\sum_{\Delta\in\mathcal{D}}
\max_{\beta\in\mathbb{Z}_{2^{16}}}
\left|
\left\{z\in\mathbb{Z}_{2^{16}} :
J_{m,K}(z+\Delta)-J_{m,K}(z)=\beta
\right\}
\right|.
\]
For the true key $K_m$, the value of $S_m(K_m)$ reflects the exact additive-differential spectrum of the inner reduced inverse cascade $Q_{m-1}^{-1}$. Wrong keys generally destroy that structure and score substantially lower.

\section{Lane-class reduction}
The autonomous second nibble of $E_i$ depends only on a 16-bit abstract signature derived from the two stage words. Concretely, the lane signature of a stage key pair $(k_0,k_1)$ at stage $i$ is
\[
\sigma_i(k_0,k_1)=
(\operatorname{nib}_2(k_0),\operatorname{nib}_2(k_2),
\operatorname{nib}_2(k_1),\operatorname{nib}_2(k_3)),
\]
where $k_2,k_3$ are the stage-local derived words used in \texttt{ENC\_Block}. This collapses the $2^{32}$ segment-key space to an average lane class of about $2^{16}$ candidates. The exact injected score above is then applied only inside the correct lane class.

\section{Verified ladder}
The file \texttt{ladder.py} contains a focused implementation of this exact injected ladder. In the \emph{state-injected} model, where the already peeled outer stage states are given, the ladder recovers all segment keys from right to left.

\begin{table}[h]
\centering
\begin{tabular}{cccc}
\toprule
Target stage & IV used & Exact rank of true key & Comment \\
\midrule
$K_8$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & $1$ & outer injected test \\
$K_7$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & $1$ & same context \\
$K_6$ & \texttt{E1A73BD303D52E6811E12AF93264B592} & $1$ & exact shortlist rescore \\
$K_5$ & \texttt{D95A1E433F62724C1FACCB1919637131} & $1$ & exact shortlist rescore \\
$K_4$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & $1$ & exact shortlist rescore \\
$K_3$ & \texttt{D95A1E433F62724C1FACCB1919637131} & $1$ & exact shortlist rescore \\
$K_2$ & \texttt{D95A1E433F62724C1FACCB1919637131} & $1$ & exact shortlist rescore \\
$K_1$ & \texttt{D95A1E433F62724C1FACCB1919637131} & $1$ & perfect linear residual \\
\bottomrule
\end{tabular}
\caption{Verified exact right-recursive ladder in the state-injected model.}
\end{table}

The most striking endpoint is stage $1$. Once the outer states are peeled, we have
\[
H_1 = B_1^{-1},
\qquad
J_{1,K_1}(z)=z-s_1.
\]
Hence every additive differential is constant:
\[
J_{1,K_1}(z+\Delta)-J_{1,K_1}(z)=\Delta
\quad\text{for all }z,\Delta.
\]
This was observed in the implementation as a perfect exact score of $6\cdot 2^{16}$ for the tested set $\mathcal{D}=\{1,2,4,8,15,16\}$.

\section{The LFSR disappears from same-reset differentials}
The Hummingbird-1 attack relies heavily on the fact that the 16-bit LFSR evolves independently from message data. The same is true here. After initialization, the update law in \texttt{main.c} is
\begin{align*}
s_2' &= s_2 + v_{12} + v_{56} + s_6,\\
s_3' &= s_3 + v_{23} + v_{34} + s_4 + s_1,\\
s_4' &= s_4 + v_{12} + v_{45} + s_8,\\
s_5' &= s_5 + v_{23} + L',\\
s_6' &= s_6 + v_{12} + v_{45} + s_7,\\
s_7' &= s_7 + v_{23} + v_{67},\\
s_8' &= s_8 + v_{45},\\
s_1' &= s_1 + s_5 + v_{34} + 2v_{23} + v_{78},
\end{align*}
where $L'=\mathrm{STEP}(L)$ is the independently updated LFSR word.

\begin{theorem}[Exact LFSR cancellation in same-reset pairs]
Consider two encryptions (or two decryptions) started from the same IV and processed for the same number of words. Let $\Delta x = x' - x$ denote additive difference modulo $2^{16}$. Then the differential state update is
\begin{align*}
\Delta s_2' &= \Delta s_2 + \Delta v_{12} + \Delta v_{56} + \Delta s_6,\\
\Delta s_3' &= \Delta s_3 + \Delta v_{23} + \Delta v_{34} + \Delta s_4 + \Delta s_1,\\
\Delta s_4' &= \Delta s_4 + \Delta v_{12} + \Delta v_{45} + \Delta s_8,\\
\Delta s_5' &= \Delta s_5 + \Delta v_{23},\\
\Delta s_6' &= \Delta s_6 + \Delta v_{12} + \Delta v_{45} + \Delta s_7,\\
\Delta s_7' &= \Delta s_7 + \Delta v_{23} + \Delta v_{67},\\
\Delta s_8' &= \Delta s_8 + \Delta v_{45},\\
\Delta s_1' &= \Delta s_1 + \Delta s_5 + \Delta v_{34} + 2\Delta v_{23} + \Delta v_{78}.
\end{align*}
In particular, the LFSR contributes no term at all to the same-reset differential recurrence.
\end{theorem}

\begin{proof}
The two executions have the same initial LFSR value because they use the same IV, and the LFSR evolves independently of message words. Hence the same $L'$ is added to both copies of $s_5$. Subtracting the two update laws therefore cancels the LFSR term exactly.
\end{proof}

This is the exact analogue of the Hummingbird-1 observation that the independent LFSR is not a meaningful barrier to differential cryptanalysis.

\section{A four-word exact outer-even core}
The previous theorem says that the LFSR disappears in differential space, but we can go further and isolate an even smaller exact state basis than the five-word core used in the earlier draft.

\begin{definition}[Outer-even core]
For round $t$, define the four-word outer-even core
\[
Y_t=(s_{2,t},s_{4,t},s_{6,t},s_{8,t}).
\]
\end{definition}

\begin{theorem}[Exact full-state reconstruction from two outer-even values]
Let $p_t$ be the round-$t$ plaintext word, let $Y_t=(s_2,s_4,s_6,s_8)$, and let $Y_{t+1}=(s_2',s_4',s_6',s_8')$. Then the full round-$t$ state
\[
S_t=(s_1,s_2,s_3,s_4,s_5,s_6,s_7,s_8)
\]
is determined exactly by
\begin{align*}
v_{45} &= s_8' - s_8,\\
v_{12} &= s_4' - s_4 - s_8',\\
v_{56} &= s_2' - s_2 - s_6 - v_{12},\\
s_1 &= D_1(v_{12}) - p_t,\\
s_5 &= D_5(v_{56}) - v_{45},\\
s_7 &= s_6' - s_6 - v_{12} - v_{45},\\
v_{23} &= E_2(v_{12}+s_2),\\
v_{34} &= D_4(v_{45}) - s_4,\\
s_3 &= D_3(v_{34}) - v_{23},
\end{align*}
where $D_i=E_i^{-1}$ is the stage-local inverse block.
\end{theorem}

\begin{proof}
The update law gives $v_{45}=s_8'-s_8$ immediately and hence
\[
v_{12}=s_4'-s_4-v_{45}-s_8 = s_4'-s_4-s_8'.
\]
Then
\[
v_{56}=s_2'-s_2-s_6-v_{12}.
\]
Stage $1$ is $v_{12}=E_1(p_t+s_1)$, so $s_1=D_1(v_{12})-p_t$. Stage $5$ is $v_{56}=E_5(v_{45}+s_5)$, so $s_5=D_5(v_{56})-v_{45}$. The $s_6$ update gives
\[
s_6'=s_6+v_{12}+v_{45}+s_7,
\]
hence $s_7=s_6'-s_6-v_{12}-v_{45}$. Since stage $2$ is $v_{23}=E_2(v_{12}+s_2)$, $v_{23}$ is now known exactly. Stage $4$ gives $v_{34}=D_4(v_{45})-s_4$, and stage $3$ then gives $s_3=D_3(v_{34})-v_{23}$.
\end{proof}

\begin{corollary}[The old five-word core is redundant]
From $(Y_t,Y_{t+1},p_t)$ one also recovers
\[
X_t=(s_{2,t},s_{4,t},s_{6,t},s_{7,t},s_{8,t}),
\]
so the previous five-word core is strictly dominated by the outer-even four-word core.
\end{corollary}

\begin{proof}
The theorem recovers $s_{7,t}$ exactly, while the other four coordinates of $X_t$ are already present in $Y_t$.
\end{proof}

\begin{corollary}[Exact LFSR recovery from two adjacent outer-even pairs]
Given $Y_t$, $Y_{t+1}$, $Y_{t+2}$ and plaintext words $p_t,p_{t+1}$, one recovers the stepped LFSR word exactly:
\[
L_{t+1}=s_{5,t+1}-s_{5,t}-v_{23,t}.
\]
\end{corollary}

\begin{proof}
Apply the theorem to $(Y_t,Y_{t+1},p_t)$ and $(Y_{t+1},Y_{t+2},p_{t+1})$ to recover $s_{5,t}$, $s_{5,t+1}$ and $v_{23,t}$ exactly. The update equation
\[
s_{5,t+1}=s_{5,t}+v_{23,t}+L_{t+1}
\]
then yields the stated formula.
\end{proof}

In fact, among same-index state subsets reused at both rounds, the set $\{2,4,6,8\}$ is the unique size-$4$ subset from which the full current state is recovered by sequential elimination of the exact update equations. This was verified exhaustively in the companion implementation.

This shows that the independent 16-bit LFSR does not enlarge the true algebraic difficulty of the state problem. Once the outer key ladder is in hand, public state recovery need only target the four-word core $Y_t$ across adjacent contexts; the remaining four state words and the stepped LFSR word are explicit consequences.

\section{A suffix-bridge reduction}
For the current attack, the most useful state relation is slightly different from the outer-even theorem above. Once the right suffix keys have been recovered, a \emph{single} observed bridge transition already collapses most of the remaining state exactly.

\begin{definition}[Suffix bridge data]
For round $t$, define the current suffix
\[
T_t=(s_{5,t},s_{6,t},s_{7,t},s_{8,t}),
\]
the next three-word suffix tail
\[
U_{t+1}=(s_{6,t+1},s_{7,t+1},s_{8,t+1}),
\]
and the bridge transcript $(p_t,c_t)$ of the round-$t$ plaintext and ciphertext.
\end{definition}

\begin{theorem}[Exact suffix-bridge extraction]
Assume $K_5,K_6,K_7,K_8$ are known. From $(T_t,U_{t+1},p_t,c_t)$ one recovers exactly
\begin{align*}
v_{78} &= D_8(c_t)-s_8,\\
v_{67} &= D_7(v_{78})-s_7,\\
v_{56} &= D_6(v_{67})-s_6,\\
v_{45} &= s_8'-s_8,\\
s_5 &= D_5(v_{56})-v_{45},\\
v_{23} &= s_7'-s_7-v_{67},\\
v_{12} &= s_6'-s_6-v_{45}-s_7.
\end{align*}
\end{theorem}

\begin{proof}
The rightmost three equations are immediate from the decryption chain:
\[
v_{78}=D_8(c_t)-s_8,\qquad
v_{67}=D_7(v_{78})-s_7,\qquad
v_{56}=D_6(v_{67})-s_6.
\]
The state update law gives
\[
s_8'=s_8+v_{45},
\]
hence $v_{45}=s_8'-s_8$. Since stage $5$ is $v_{56}=E_5(v_{45}+s_5)$, we obtain
\[
s_5=D_5(v_{56})-v_{45}.
\]
The update laws
\[
s_7'=s_7+v_{23}+v_{67},\qquad
s_6'=s_6+v_{12}+v_{45}+s_7
\]
then give the stated formulas for $v_{23}$ and $v_{12}$.
\end{proof}

\begin{corollary}[Immediate recovery of $s_1$ and $s_2$]
If $K_1$ and $K_2$ are also known, then the same bridge data determines
\[
s_1=D_1(v_{12})-p_t,\qquad
s_2=D_2(v_{23})-v_{12}.
\]
\end{corollary}

\begin{proof}
Stage $1$ is $v_{12}=E_1(p_t+s_1)$, so $s_1=D_1(v_{12})-p_t$. Stage $2$ is $v_{23}=E_2(v_{12}+s_2)$, so $s_2=D_2(v_{23})-v_{12}$.
\end{proof}

\begin{corollary}[One-word left-state family]
Assume $K_1,K_2,K_3,K_4$ are known in addition to the suffix keys. Then the bridge data determines the current left half up to a single 16-bit word:
\[
s_4' = s_4 + v_{12} + s_8',
\]
and for every candidate $s_4$,
\[
s_3 = D_3(D_4(v_{45})-s_4)-v_{23}.
\]
Equivalently, fixing either $s_4$ or $s_4'$ collapses the family to a unique full round state.
\end{corollary}

\begin{proof}
The update law for $s_4$ is
\[
s_4' = s_4 + v_{12} + v_{45} + s_8 = s_4 + v_{12} + s_8'.
\]
Stage $4$ gives $v_{34}=D_4(v_{45})-s_4$, and stage $3$ then yields
\[
s_3=D_3(v_{34})-v_{23}=D_3(D_4(v_{45})-s_4)-v_{23}.
\]
Thus all unknown coordinates are explicit functions of the single free word $s_4$ (or equivalently $s_4'$).
\end{proof}

In the companion implementation, this reduced family is also checked with a small SMT solver: without fixing $s_4'$ the solver finds multiple models differing only in the free 16-bit bridge word, while fixing $s_4'$ yields a unique solution.

\section{A public exact state equation}
The same ladder identity used for stage-key recovery can also be turned around and used as a \emph{public state test}.

Assume the outer right stage keys $K_{m-1},K_m,\dots,K_8$ are known and that the already peeled outer states $s_{m+1},\dots,s_8$ are given. Let $G$ be the public inverse next-word permutation for the chosen context and define
\[
H_{m-1}^{(\tau)} =
G\circ B_8^{(s_8)}\circ \cdots \circ B_{m+1}^{(s_{m+1})}\circ B_m^{(\tau)},
\]
where
\[
B_i^{(\sigma)}(x)=E_i(x+\sigma).
\]

\begin{theorem}[Exact public state test]
For the true state word $\tau=s_m$ and the true next visible stage key $K_{m-1}$,
\[
H_{m-1}^{(s_m)}(E_{m-1}(z;K_{m-1}))
=
Q_{m-2}^{-1}(z-s_{m-1}).
\]
Hence the exact additive-differential score of the right-hand side can be used directly as a public test for $s_m$.
\end{theorem}

\begin{proof}
The already peeled outer stages cancel exactly, and for the true guess $\tau=s_m$ we obtain
\[
B_m^{-1}(E_m(z+s_m))=z.
\]
Injecting the next visible key $K_{m-1}$ then reduces the composition to the inner inverse cascade $Q_{m-2}^{-1}$ with only the next state word $s_{m-1}$ left as an input translation.
\end{proof}

Thus state recovery may proceed by the same divide-and-conquer logic as key recovery:
\begin{enumerate}[label=(\roman*)]
\item guess the current outer state word $s_m$;
\item peel it publicly into the inverse oracle;
\item inject the already known next key $K_{m-1}$;
\item select the state guess that maximizes the exact next-stage ladder score.
\end{enumerate}

In the present Python implementation a full $2^{16}$ exact state scan is still costly, so the script performs a bytewise version of this exact test: it samples all 256 candidates of one byte, exact-rescores the shortlist, and then repeats for the other byte.

\subsection*{Verified public state-byte recoveries}
The new \texttt{state-byte-scan} mode in \texttt{ladder.py} verifies this public exact state equation on the first two right-side states.

\begin{itemize}
\item For $s_8$ on IV \texttt{4703EAC61B442157747A61DDA8FDDDD3}, using the public exact $K_7$ score:
  \begin{itemize}
  \item low byte: sampled rank $2$, exact shortlist rank $1$ at the true value $69$;
  \item upper byte (with low byte fixed): sampled rank $6$, exact shortlist rank $1$ at the true value $2B$.
  \end{itemize}
\item For $s_7$ on IV \texttt{E1A73BD303D52E6811E12AF93264B592}, with the correct outer state $s_8=\texttt{1ECC}$ peeled and using the public exact $K_6$ score:
  \begin{itemize}
  \item low byte: sampled rank $1$, exact shortlist rank $1$ at the true value $02$;
  \item upper byte (with low byte fixed): sampled shortlist contains the true value, and exact shortlist rank $1$ at the true value $D5$.
  \end{itemize}
\end{itemize}

These experiments are not yet a full public recursive state recovery, but they show that the exact ladder signal remains usable on the state side once the next stage key is known.

\section{An exact low-byte state primitive}
There is a second exact state observable that is cheaper than the full next-stage ladder score and, unlike the high-byte problem, does not depend on fixing the output translation gauge.

Let $F_m$ be a public forward reduced table at stage $m$ after all outer stages $m+1,\dots,8$ have been peeled exactly. For the true stage key $K_m$, define the peeled table
\[
W_m(x)=D_m(F_m(x)).
\]
For the true state word $s_m = 256h + \ell$ this has the form
\[
W_m(x)=T_{m-1}(x)+256h+\ell,
\]
where $T_{m-1}$ is the remaining inner cascade.

For a fixed row $r\in\{0,\dots,255\}$, define
\[
\sigma_r(b)=\left|\left\{\operatorname{hi}\bigl(W_m(256r+\lambda)-b\bigr): \lambda\in\{0,\dots,255\}\right\}\right|,
\]
and the all-row score
\[
\Sigma(b)=\sum_{r=0}^{255}\sigma_r(b).
\]

Subtracting a candidate low byte $b$ changes the upper byte only through the carry bit from the low byte. Hence $\Sigma(b)$ is an exact deterministic carry-profile score for the unknown low byte $\ell$.

\subsection*{Verified exact low-byte recoveries}
The new \texttt{low-byte-scan} mode in \texttt{ladder.py} verifies that, on the tested exact right-recursive contexts, the true low byte is the unique minimizer of $\Sigma(b)$.

\begin{center}
\begin{tabular}{cccc}
\toprule
stage & IV & true word & exact low-byte rank \\
\midrule
$8$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & \texttt{2B69} & $1$ \\
$7$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & \texttt{C29E} & $1$ \\
$6$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & \texttt{8E72} & $1$ \\
$4$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{D547} & $1$ \\
$2$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{267C} & $1$ \\
$1$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{599F} & $1$ \\
\bottomrule
\end{tabular}
\end{center}

Representative scores were:
\begin{itemize}
\item stage $8$: $\Sigma(\texttt{69})=3531$, next best $\Sigma(\texttt{68})=3763$;
\item stage $7$: $\Sigma(\texttt{9E})=2413$, next best $\Sigma(\texttt{9D})=2657$;
\item stage $6$: $\Sigma(\texttt{72})=1654$, next best $\Sigma(\texttt{71})=1900$;
\item stage $2$: $\Sigma(\texttt{7C})=512$, next best $\Sigma(\texttt{7B})=768$;
\item stage $1$: $\Sigma(\texttt{9F})=256$, all next-best scores start at $512$.
\end{itemize}

So the current exact public state picture is now asymmetric:
\begin{itemize}
\item the low byte is directly exposed by a deterministic all-row carry-support score;
\item the remaining part is the state high byte, which can now also be exposed exactly once the next visible stage key is known.
\end{itemize}

\section{An exact high-byte and next-key support primitive}
The low-byte primitive above removes the carry ambiguity from the current stage translation. The remaining hidden quantity is the current stage high byte. The key point is that, after the current low byte is removed, the next peeled stage sees only an \emph{upper-byte} translation before its own byte-triangular permutation.

Let $F_m$ again be the forward reduced table at stage $m$, and let
\[
W_m(x)=D_m(F_m(x)).
\]
Write the true state as
\[
s_m = 256h_m + \ell_m.
\]
After recovering $\ell_m$ exactly, define the low-aligned table
\[
\widetilde{W}_m(x)=W_m(x)-\ell_m.
\]

For a candidate next key $K$ for stage $m-1$ and a candidate current high byte $h$, define for each upper-input row $r\in\{0,\dots,255\}$ the minimized support
\[
\rho_r(K,h)=
\min_{b\in\{0,\dots,255\}}
\left|
\left\{
\operatorname{hi}\!\left(
D_K\!\bigl(\widetilde{W}_m(256r+\lambda)-256h\bigr)-b
\right)
:\lambda\in\{0,\dots,255\}
\right\}
\right|.
\]
Then define the total support score
\[
\mathcal{H}_m(K,h)=\sum_{r=0}^{255}\rho_r(K,h).
\]

For the true pair $(K_{m-1},h_m)$, subtracting $256h_m$ removes the remaining translation of the current stage. Applying the correct next-stage inverse yields a byte-triangular reduced table again, and the minimization over $b$ removes the next stage's unknown low-byte translation row by row. Thus $\mathcal{H}_m(K,h)$ is an exact branchless score for the joint pair ``current high byte and next stage key.''

For stage $m=2$, the next stage is the first stage, and the same phenomenon becomes a constancy score. After the full current translation is removed, define
\[
\mathcal{C}_1(K,h)=
\max_{\delta\in\mathbb{Z}_{2^{16}}}
\left|
\left\{
x\in\mathbb{Z}_{2^{16}}:
D_K\!\bigl(W_2(x)-(\ell_2+256h)\bigr)-x=\delta
\right\}
\right|.
\]
For the true $(K_1,h_2)$, the peeled stage-$1$ table differs from the identity by the constant translation $s_1$, so $\mathcal{C}_1$ is maximized exactly at the true high byte and the true first-stage key.

\subsection*{Verified exact high-byte recoveries}
The new \texttt{high-byte-scan} mode in \texttt{ladder.py} validates this exact high-byte finish on the target instance.

\begin{center}
\begin{tabular}{ccccc}
\toprule
stage & IV & true word & criterion & exact high-byte rank \\
\midrule
$8$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & \texttt{2B69} & $\mathcal{H}_8(K_7,h)$ & $1$ \\
$7$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & \texttt{C29E} & $\mathcal{H}_7(K_6,h)$ & $1$ \\
$6$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{59DD} & $\mathcal{H}_6(K_5,h)$ & $1$ \\
$4$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{D547} & $\mathcal{H}_4(K_3,h)$ & $1$ \\
$2$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{267C} & $\mathcal{C}_1(K_1,h)$ & $1$ \\
\bottomrule
\end{tabular}
\end{center}

Representative exact scores were:
\begin{itemize}
\item stage $8$: the true high byte \texttt{2B} scored $136$, next best score $174$;
\item stage $7$: the true high byte \texttt{C2} scored $100$, next best score $119$;
\item stage $6$: the true high byte \texttt{59} scored $237$, next best score $319$;
\item stage $4$: the true high byte \texttt{D5} scored $64$, next best score $101$;
\item stage $2$: the true high byte \texttt{26} achieved the perfect constancy score $2^{16}=65536$, while the next best scores were only $468$.
\end{itemize}

\subsection*{Verified joint next-key ranking on sampled lane classes}
The same score also ranks the next visible key correctly on sampled same-lane candidate bundles.

\begin{itemize}
\item stage $8 \rightarrow 7$, IV \texttt{4703EAC61B442157747A61DDA8FDDDD3}: on an $8$-pair sampled stage-$7$ lane class, the true $K_7=(\texttt{8D1E},\texttt{9DF5})$ ranked $1$ and the minimizing high byte was the true value \texttt{2B};
\item stage $7 \rightarrow 6$, same IV with $s_8=\texttt{2B69}$ peeled: on an $8$-pair sampled stage-$6$ lane class, the true $K_6=(\texttt{F44A},\texttt{233E})$ ranked $1$ and the minimizing high byte was the true value \texttt{C2}.
\end{itemize}

So the remaining public state/key problem is now much more structured than before:
\begin{itemize}
\item recover the current low byte exactly by the all-row carry/support minimum;
\item recover the current high byte and the next key jointly by the exact branchless support score $\mathcal{H}_m$;
\item use the stage-$1$ constancy score $\mathcal{C}_1$ at the bottom of the recursion.
\end{itemize}

\subsection*{A Hummingbird-style exact one-step lookahead}
The next natural step, inspired by the multi-step exact equations in the Hummingbird attacks, is to use the recovered current state word as a \emph{recursive exact consistency check} on the next stage.

For a candidate next key $K$ at stage $m-1$, let
\[
\widehat{h}_m(K)=\arg\min_h \mathcal{H}_m(K,h),
\qquad
\widehat{s}_m(K)=\ell_m + 256\widehat{h}_m(K),
\]
where $\ell_m$ is the exact low byte recovered by the carry/support primitive. Define the candidate next peeled table
\[
\widehat{W}_{m-1}^{(K)}(x)=D_K\!\bigl(W_m(x)-\widehat{s}_m(K)\bigr).
\]
If $K=K_{m-1}$, then
\[
\widehat{W}_{m-1}^{(K_{m-1})}(x)=T_{m-2}(x)+s_{m-1},
\]
so the exact low-byte primitive applies again to $\widehat{W}_{m-1}^{(K_{m-1})}$ and must recover the true next low byte $\ell_{m-1}$.

This gives a second exact criterion on the same candidate:
\begin{enumerate}[label=(\roman*)]
\item minimize the current high-byte support score $\mathcal{H}_m(K,h)$;
\item peel with the resulting $\widehat{s}_m(K)$;
\item require the next-stage low-byte support score to behave like a genuine reduced stage.
\end{enumerate}

Thus the attack is no longer a one-layer exact test. It is a short exact equation cascade:
\[
K_{m-1}\;\Longrightarrow\;\widehat{h}_m(K)\;\Longrightarrow\;\widehat{W}_{m-1}^{(K)}
\;\Longrightarrow\;\text{exact next low byte}.
\]

\subsection*{Verified recursive exact lookahead}
The new \texttt{recursive-pair-scan} mode in \texttt{ladder.py} validates this one-step exact lookahead on sampled same-lane candidate bundles.

\begin{center}
\begin{tabular}{cccccc}
\toprule
current stage & IV & true next key & true next low & exact pair rank & best wrong next low \\
\midrule
$8 \rightarrow 7$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & \texttt{8D1E,9DF5} & \texttt{9E} & $1$ & \texttt{00} \\
$7 \rightarrow 6$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & \texttt{F44A,233E} & \texttt{72} & $1$ & \texttt{00} \\
$6 \rightarrow 5$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{50C5,12DB} & \texttt{DB} & $1$ & \texttt{00} \\
$4 \rightarrow 3$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{02DD,3972} & \texttt{62} & $1$ & \texttt{00} \\
\bottomrule
\end{tabular}
\end{center}

Representative measured tuples were:
\begin{itemize}
\item stage $8 \rightarrow 7$: true pair $(136,\texttt{2B},2413,244)$ versus wrong pairs such as $(171,\texttt{A0},3531,247)$;
\item stage $7 \rightarrow 6$: true pair $(100,\texttt{C2},1654,246)$ versus wrong pairs such as $(109,\texttt{95},2413,244)$;
\item stage $6 \rightarrow 5$: true pair $(237,\texttt{59},3697,236)$ versus wrong pairs such as $(316,\texttt{9E},5403,229)$;
\item stage $4 \rightarrow 3$: true pair $(64,\texttt{D5},1024,254)$ versus wrong pairs such as $(96,\texttt{42},2017,244)$.
\end{itemize}
Here each tuple is
\[
(\text{current high-score},\ \widehat{h}_m,\ \text{next low support score},\ \text{next low gap}).
\]

This is the cleanest recursive public filter obtained so far. Wrong candidates do not merely score worse at the current stage; after the candidate peel they fail to produce the exact next-stage low-byte structure and typically collapse to the generic low-byte winner \texttt{00}. That is much closer to the exact multi-step logic used in the Hummingbird cryptanalysis than the earlier one-layer public scans.

\section{An exact upper-transport law on the reduced inverse chain}
The previous sections worked with the right-recursive forward peel. There is a second exact recursion law on the true reduced inverse chain that is even closer to the ``prepared suffix'' logic in Hummingbird-style attacks.

Let
\[
T_i = B_i^{-1}\circ B_{i+1}^{-1}\circ\cdots\circ B_8^{-1},
\qquad i=1,\dots,8,
\]
so that
\[
T_i(x)=T_{i+1}(D_i(x)-s_i).
\]
Write
\[
s_i = 256h_i + \ell_i.
\]
For a row index $r\in\{0,\dots,255\}$ define the low-corrected upper support of the current reduced table by
\[
\mathcal{S}_i(r)=
\left\{
\operatorname{hi}\bigl(T_i(256r+\lambda)+\ell_i\bigr)
:\lambda\in\{0,\dots,255\}
\right\},
\]
and define the raw next-stage upper support by
\[
\mathcal{R}_{i+1}(r)=
\left\{
\operatorname{hi}\bigl(T_{i+1}(256r+\lambda)\bigr)
:\lambda\in\{0,\dots,255\}
\right\}.
\]

Let $U_i$ be the upper-byte map of the inverse stage block:
\[
U_i(r)=\operatorname{hi}(D_i(256r)).
\]

\begin{theorem}[Exact upper-transport law]
For every stage $i\in\{1,\dots,7\}$, the multiset of low-corrected upper supports of $T_i$ is exactly the stage-$i$ upper map applied to the raw next-stage supports, followed by a global additive shift $-h_i$:
\[
\bigl\{\mathcal{S}_i(r):r\in\mathbb{Z}_{256}\bigr\}
=
\bigl\{U_i(\mathcal{R}_{i+1}(r))-h_i:r\in\mathbb{Z}_{256}\bigr\}.
\]
Equivalently, the row order may permute, but the full row-support multiset transports exactly through the current inverse stage.
\end{theorem}

\begin{proof}
For fixed $r$ and $\lambda$, write
\[
D_i(256r+\lambda)=256u+\mu.
\]
Because $D_i$ is byte-triangular, the upper byte $u$ depends only on $r$. Subtracting $s_i=256h_i+\ell_i$ gives
\[
D_i(256r+\lambda)-s_i = 256(u-h_i-c(\lambda)) + (\mu-\ell_i),
\]
where $c(\lambda)\in\{0,1\}$ is the borrow from subtracting the low byte. Adding back $\ell_i$ to the \emph{output} of $T_i$ removes the low-byte gauge and leaves only the upper translation by $-h_i$ after transport through $T_{i+1}$. Since $u$ is row-constant and $T_{i+1}$ contributes only the raw next-stage support set seen at that transported row, the current row support is exactly $U_i$ applied to the corresponding next-stage support, shifted by $-h_i$. Taking multisets over all rows removes the row permutation.
\end{proof}

The theorem has two concrete attack consequences.
\begin{enumerate}[label=(\roman*)]
\item Once the next reduced table $T_{i+1}$ is known, the current pair $(K_i,s_i)$ can be tested by \emph{exact} multiset equality rather than by a differential score.
\item The unknown current high byte $h_i$ is no longer a nuisance variable; it is exactly the unique shift that makes the transported multiset match.
\end{enumerate}

\subsection*{Verified exact transport validation}
The new \texttt{transport-validate} mode in \texttt{ladder.py} checks this theorem directly on fresh random contexts. On two independent random IVs, every stage $1$ through $7$ matched exactly:
\[
14/14
\]
full upper-support multiset matches and
\[
14/14
\]
canonical multiset matches.

\subsection*{Verified exact pair recovery from the next reduced table}
The stronger point is that the transport law is selective enough to identify the true current stage pair on sampled same-lane bundles.

\begin{center}
\begin{tabular}{cccccc}
\toprule
stage & IV & true pair & true state & sampled bundle & exact transport rank \\
\midrule
$7$ & \texttt{4703EAC61B442157747A61DDA8FDDDD3} & \texttt{8D1E,9DF5} & \texttt{C29E} & $8$ pairs & $1$ \\
$6$ & \texttt{E1A73BD303D52E6811E12AF93264B592} & \texttt{F44A,233E} & \texttt{CA2F} & $8$ pairs & $1$ \\
$5$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{50C5,12DB} & \texttt{A1DB} & $8$ pairs & $1$ \\
$1$ & \texttt{D95A1E433F62724C1FACCB1919637131} & \texttt{E8B9,B733} & \texttt{599F} & $8$ pairs & $1$ \\
\bottomrule
\end{tabular}
\end{center}

In each checked bundle, the true pair was the \emph{only} candidate with a full multiset match, and the recovered shift was the true value $-h_i \bmod 256$.

So there are now two exact suffix mechanisms available after the bootstrap:
\begin{itemize}
\item the injected-differential ladder of Sections 5--8;
\item the reduced-inverse transport law of this section.
\end{itemize}
Both say the same thing in different coordinates: once the outer gauge is fixed, the remainder of the recursion is exact algebra rather than residual brute force.

\section{An exact related-state trail}
The previous sections attacked a single matched context. We now describe a second, independent exact algebraic primitive that is much closer in spirit to the original Hummingbird-1 cryptanalysis: an exact related-state trail that converts a chosen-IV relation into deterministic one-word behaviour.

\begin{theorem}[Exact one-word related-state family]
Let two online SEPAR states differ only in the first word:
\[
\Delta(s_1,s_2,s_3,s_4,s_5,s_6,s_7,s_8)=(\delta,0,0,0,0,0,0,0).
\]
Then:
\begin{enumerate}[label=(\roman*)]
\item in encryption, if the plaintext difference is $\Delta p=-\delta$, the ciphertext difference is exactly $0$ and the post-round state difference is
\[
(\delta,0,\delta,0,0,0,0,0);
\]
\item in decryption, if the ciphertext difference is $0$, the plaintext difference is exactly $\Delta p=-\delta$ and the post-round state difference is again
\[
(\delta,0,\delta,0,0,0,0,0).
\]
\end{enumerate}
\end{theorem}

\begin{proof}
For encryption, the first stage input is $p+s_1$. Under the related pair
\[
(p,s_1)\mapsto (p-\delta,s_1+\delta),
\]
this quantity is unchanged, so the stage-$1$ output $v_{12}$ is identical in both executions. Since $s_2,\dots,s_8$ are equal initially, every later stage input and every later stage output is also identical, so the ciphertext words are equal. Substituting $\Delta v_{12}=\cdots=\Delta v_{78}=0$ into the update law gives
\[
\Delta s_1'=\delta,\qquad
\Delta s_3'=\delta,
\]
and every other post-round state difference is $0$.

For decryption, the same-ciphertext condition implies that the inverse stages $8,7,\dots,2$ all receive identical inputs because $s_2,\dots,s_8$ are equal. Hence $v_{78},v_{67},\dots,v_{12}$ are identical in both executions. The only difference appears at the final subtraction
\[
p=D_1(v_{12})-s_1,
\]
which immediately gives $\Delta p=-\delta$. The post-round state difference follows from the same update substitution as above.
\end{proof}

This family was validated exhaustively over random states in \texttt{ladder.py} for several values of $\delta$, including $\delta=\texttt{8000}$ and $\delta=\texttt{0001}$.

\subsection*{Exact second-word exposure}
The one-word trail above does not iterate unchanged, but it produces a \emph{deterministic} next-round state:
\[
\Delta S_1=(\delta,0,\delta,0,0,0,0,0).
\]
That already exposes a pure inner-stage differential on the next word.

\paragraph{Encryption.}
Run a second encryption word under the same plaintext difference $\Delta p=-\delta$. Then
\[
\Delta(p+s_1)=0,\qquad \Delta(v_{12}+s_2)=0,
\]
so
\[
\Delta v_{12}=0,\qquad \Delta v_{23}=0.
\]
The first nonzero difference therefore appears at stage $3$:
\[
\Delta(v_{23}+s_3)=\delta.
\]
Thus the second word exposes an exact stage-$3$ block differential
\[
v_{34}'-v_{34}=E_3(x+\delta)-E_3(x),
\qquad
x=v_{23}+s_3.
\]

\paragraph{Decryption.}
Run a second decryption word under the same ciphertext in both instances. Since only $s_1$ and $s_3$ differ,
\[
\Delta v_{78}=\Delta v_{67}=\Delta v_{56}=\Delta v_{45}=\Delta v_{34}=0,
\]
and the first nonzero difference appears at stage $3$ inverse:
\[
v_{23}'-v_{23}=-\delta.
\]
Hence the next stage-$2$ inverse input is known exactly:
\[
\Delta(v_{23})=-\delta,
\qquad
v_{12}'-v_{12}=D_2(x-\delta)-D_2(x),
\qquad
x=v_{23}.
\]

So once a suitable chosen-IV pair has been found, the \emph{second} word gives an exact single-stage differential exposure: stage $3$ on the encryption side and stage $2$ on the decryption side.

\paragraph{Current status.}
The exact trail itself is solved and implemented. The remaining open step is the Hummingbird-style chosen-IV pair search: one needs two IVs whose post-initialization states satisfy the required one-word relation
\[
\Delta S_0=(\delta,0,0,0,0,0,0,0).
\]
The companion script now includes a public fingerprint search harness for such pairs, but a practical pair-finding algorithm is still open.

\section{Complexity in the injected-state model}
For one matched context, the data cost is the cost of building the next-word codebook:
\[
2^{16}
\]
chosen words for the forward table and the same order of work for the inverse table if both are built explicitly.

Offline, each exact stage scan costs
\[
|\mathcal{C}_m|\cdot |\mathcal{D}|\cdot 2^{16}
\]
block evaluations, where $\mathcal{C}_m$ is the scanned lane class for stage $m$. In the present implementation the lane class is sampled first and then the best shortlist is exact-rescored. The empirical takeaway is not the precise constant, but the stronger structural statement:

\begin{quote}
Once the outer right states are provided, the broken implementation admits an exact right-recursive ladder that recovers $K_8,\dots,K_1$ stage by stage.
\end{quote}

\section{Toward full public key recovery}
The ladder above removes the stage-key problem from the list of open issues. The remaining obstacle is public recovery of the current outer state word $s_m$ at each step.

The exact recursion suggests the right public architecture:
\begin{enumerate}[label=(\arabic*)]
\item recover the current outer right key/state pair $(K_m,s_m)$;
\item form $H_m$ publicly;
\item use the injected exact score to recover $K_{m-1}$ without depending on $s_{m-1}$;
\item recurse.
\end{enumerate}

The new state algebra refines this picture further: one does not need to recover the full
\[
(s_1,\dots,s_8,L)
\]
at each stage. It is enough to recover the four-word outer-even core $Y_t=(s_2,s_4,s_6,s_8)$ on adjacent contexts, because $(s_1,s_3,s_5,s_7)$ and the stepped LFSR word then follow by exact formulas.

The current state of the attack is therefore:
\begin{itemize}
\item \textbf{Solved exactly:} right-recursive stage-key recovery under injected outer states;
\item \textbf{Reduced exactly:} the LFSR/state problem collapses to recovery of the four-word outer-even core on adjacent contexts;
\item \textbf{Solved publicly at the byte level:} the current state low byte is exposed exactly by the all-row carry-support minimum, and the remaining high byte is exposed exactly by $\mathcal{H}_m$ or $\mathcal{C}_1$ once the next visible key candidate is supplied;
\item \textbf{Partially solved publicly at the key/state level:} on sampled same-lane bundles, $\mathcal{H}_m$ ranks the true next key first while returning the true current high byte;
\item \textbf{Still open:} turning that exact byte/key algebra into a fully public end-to-end recursion without truth-guided lane-class injection.
\end{itemize}

\section{Reproducing the ladder}
The companion script \texttt{ladder.py} provides the following core modes.
\begin{itemize}
\item \texttt{demo}: runs the verified ladder on known good contexts;
\item \texttt{scan}: evaluates a single stage and IV;
\item \texttt{search}: searches random IVs for contexts where the sampled injected score is sharp.
\item \texttt{outer4-demo}: demonstrates exact full-state and LFSR reconstruction from two adjacent outer-even pairs.
\item \texttt{outer4-validate}: validates the exact same-reset differential recurrence and the outer-even reconstruction formulas over random contexts.
\item \texttt{core-demo}: retains the earlier five-word-core reconstruction for comparison.
\item \texttt{core-validate}: retains validation of the earlier five-word-core formulas.
\item \texttt{trail-validate}: validates the exact related-state family and the exact second-word stage exposures described above.
\item \texttt{iv-pair-search}: performs a public chosen-IV fingerprint search for candidate related-IV pairs compatible with the exact trail.
\item \texttt{state-byte-scan}: demonstrates the public exact state-byte test described above.
\item \texttt{low-byte-scan}: demonstrates the exact all-row carry/support recovery of the current state low byte.
\item \texttt{high-byte-scan}: demonstrates the exact branchless support / constancy recovery of the current state high byte and, on sampled bundles, the next visible key.
\item \texttt{recursive-pair-scan}: adds a one-step exact lookahead and ranks the next visible key by whether the resulting candidate peel also yields the correct next-stage low-byte structure.
\item \texttt{transport-validate}: validates the exact upper-transport law across stages $1$ through $7$ on fresh random contexts.
\item \texttt{transport-pair-scan}: uses the true next reduced table to show that the transport law identifies the current stage pair and the current high-byte shift exactly on sampled same-lane bundles.
\end{itemize}

Typical commands are:
\begin{verbatim}
python ladder.py --mode scan --iv 4703EAC61B442157747A61DDA8FDDDD3 \
  --target-right 8 --class-limit 64 --sample-step 2048 --sample-count 32 \
  --exact-top 2

python ladder.py --mode scan --iv D95A1E433F62724C1FACCB1919637131 \
  --target-right 1 --class-limit 64 --sample-step 2048 --sample-count 32 \
  --exact-top 8

python ladder.py --mode outer4-demo --iv D95A1E433F62724C1FACCB1919637131 \
  --words 1234,5678

python ladder.py --mode outer4-validate --samples 64 --rounds 5 --seed 11

python ladder.py --mode trail-validate --delta 8000 --samples 128 --seed 5

python ladder.py --mode iv-pair-search --delta 8000 --trials 4096 \
  --probe-words 0000,1111 --truth-check

python ladder.py --mode state-byte-scan --iv 4703EAC61B442157747A61DDA8FDDDD3 \
  --scan-stage 8 --scan-part low --sample-step 257 --sample-count 256 \
  --exact-top 4 --diffs 0001,0002,0004,0008

python ladder.py --mode state-byte-scan --iv E1A73BD303D52E6811E12AF93264B592 \
  --scan-stage 7 --scan-part low --known-right-states 8=1ECC \
  --sample-step 1024 --sample-count 64 --exact-top 4 \
  --diffs 0001,0002,0004,0008

python ladder.py --mode low-byte-scan --iv 4703EAC61B442157747A61DDA8FDDDD3 \
  --target-right 8 --top 8

python ladder.py --mode high-byte-scan --iv 4703EAC61B442157747A61DDA8FDDDD3 \
  --target-right 8 --candidate-next-pair 8D1E,9DF5 --row-step 16 --top 8

python ladder.py --mode high-byte-scan --iv 4703EAC61B442157747A61DDA8FDDDD3 \
  --target-right 7 --known-right-states 8=2B69 --class-limit 8 \
  --class-seed 1 --row-step 16 --top 8

python ladder.py --mode recursive-pair-scan --iv 4703EAC61B442157747A61DDA8FDDDD3 \
  --target-right 8 --class-limit 8 --class-seed 1 --row-step 16 --top 8

python ladder.py --mode transport-validate --samples 2 --seed 11

python ladder.py --mode transport-pair-scan --target-right 7 --class-limit 8 \
  --class-seed 1 --top 8
\end{verbatim}

\section{Conclusion}
The broken SEPAR implementation yields to an exact Hummingbird-style divide-and-conquer attack on the right side. The essential observation is not merely that differentials are ``strong,'' but that after the outer right peel one may inject a candidate block permutation directly into the reduced inverse oracle. For the true key, the current stage state cancels \emph{algebraically} from the additive differential spectrum. This gives a genuine exact recursive ladder.

The remaining public problem is also narrower than it first appears. The 16-bit LFSR disappears exactly from same-reset differential recurrences, the current state low byte is exposed by a deterministic row-support minimum, the current state high byte is exposed by an exact branchless next-stage support score once a next-key candidate is supplied, and the reduced inverse chain now satisfies an exact upper-transport law that makes the post-bootstrap suffix deterministic. Moreover, the full round state is determined by two consecutive values of the four-word outer-even core $(s_2,s_4,s_6,s_8)$. Thus the unresolved part of the attack is no longer ``defeat the LFSR'' or ``solve the inner recursion,'' but rather ``find a fully public outer bootstrap that fixes the gauge and exposes the first exact reduced chain.'' That is a substantially smaller and cleaner target for the next pass.

\end{document}
